### 2024-11-27 paper/paper.qmd
changed name

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Dennis Netchitailo
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-11-27 britain-bombing-ww2/src/add-location-data.R
add downloaded data

@@ -1,0 +1,36 @@
+
+### Load packages ---------------------------------------------------------------------------------
+library(tidyverse)
+library(readxl)
+library(janitor)
+library(ggmap)
+
+### Load data and clean locations enough for Geocode API lookup -----------------------------------
+locations <- read_xlsx("./britain-bombing-ww2/raw/Bombing-Britain-data.xlsx",
+                       guess_max = 1e5) %>%
+  rename(location = Location,
+         country = Country) %>%
+  distinct(location, country) %>%
+  # Take only first item in composite location descriptions (multiple locations separated by +)
+  mutate(location_clean = str_remove(location, " \\+.*")) %>%
+  # Combine location and country to make lookup locations as clear as possible
+  unite(lookup_location, location_clean, country, sep = ", ", remove = FALSE) %>%
+  # Ditch unneeded fields
+  select(location, country, lookup_location)
+
+
+### Set up Google Geocode API and look up coordinates using {ggmap} -------------------------------
+### NOTE: this step can cost money $$$ (check Google pricing)
+
+# Register API key
+register_google(key = read_file("./britain-bombing-ww2/ref/keys/google-geocoding-api-key.txt"))
+
+# Look up geographic coordinates
+geocodes <- locations %>%
+  mutate(coord_data = map(lookup_location, geocode)) %>%
+  select(-lookup_location) %>%
+  unnest(coord_data)
+
+
+### Write to CSV ----------------------------------------------------------------------------------
+write_csv(geocodes, "./britain-bombing-ww2/ref/geocodes.csv")

### 2024-11-27 britain-bombing-ww2/src/clean-bombing-data.R

@@ -1,0 +1,67 @@
+
+### Load packages ---------------------------------------------------------------------------------
+library(tidyverse)
+library(readxl)
+library(janitor)
+
+
+### Load data and geocodes ------------------------------------------------------------------------
+bombings_raw <- read_xlsx("./britain-bombing-ww2/raw/Bombing-Britain-data.xlsx",
+                          guess_max = 1e5,
+                          na = c("", "N/A"))
+
+geocodes <- read_csv("./britain-bombing-ww2/ref/geocodes.csv")
+
+
+### Clean data ------------------------------------------------------------------------------------
+bombings_clean <- bombings_raw %>%
+  clean_names() %>%
+  # Add geographic coordinates (latitude and longitude)
+  left_join(geocodes, by = c("location", "country")) %>%
+  # Convert Excel dates to R date format
+  mutate(across(start_date:end_date, ~ excel_numeric_to_date(parse_number(.)))) %>%
+  # Parse numbers from casualty fields (they sometimes have values like "a few" or "unspecified")
+  mutate(across(killed:total_casualties, parse_number)) %>%
+  # Convert coordinates in the United States (clearly wrong) into missing values
+  mutate(lon = ifelse(lon < -25, NA, lon),
+         lat = ifelse(lon < -25, NA, lat)) %>%
+  # Add id field
+  mutate(bombing_id = row_number()) %>%
+  # Get rid of unnecessary fields
+  select(-volume_reference, -intelligence_summary_number, -link_to_page) %>%
+  # Re-order fields to be easier to work with
+  select(bombing_id, civil_defence_region, country, location,
+         lon, lat, start_date:time, killed:additional_notes)
+
+
+### Identify duplicate casualty records -----------------------------------------------------------
+# Identify duplicated casualty figures that are part of aggregations
+# We will create "casualty groups" that tick up by one when we conclude that the casualty data is not duplicated
+# This way, duplicated casualty figures will have the same "casualty group" ID, which we can use to prevent double-counting
+bombings_geocoded <- bombings_clean %>%
+  # The first row where one of these fields is different means a new casualty group
+  group_by(civil_defence_region, start_date, end_date, time, killed, injured, total_casualties) %>%
+  mutate(first_row = row_number() == 1) %>%
+  ungroup() %>%
+  # New casualty group when additional_notes fields doesn't include "Total" or "Entire", which are markers of aggregation
[…]

### 2024-11-27 britain-bombing-ww2/src/add-location-data.R
delete unnecessary files

@@ -1,36 +1,0 @@
-
-### Load packages ---------------------------------------------------------------------------------
-library(tidyverse)
-library(readxl)
-library(janitor)
-library(ggmap)
-
-### Load data and clean locations enough for Geocode API lookup -----------------------------------
-locations <- read_xlsx("./britain-bombing-ww2/raw/Bombing-Britain-data.xlsx",
-                       guess_max = 1e5) %>%
-  rename(location = Location,
-         country = Country) %>%
-  distinct(location, country) %>%
-  # Take only first item in composite location descriptions (multiple locations separated by +)
-  mutate(location_clean = str_remove(location, " \\+.*")) %>%
-  # Combine location and country to make lookup locations as clear as possible
-  unite(lookup_location, location_clean, country, sep = ", ", remove = FALSE) %>%
-  # Ditch unneeded fields
-  select(location, country, lookup_location)
-
-
-### Set up Google Geocode API and look up coordinates using {ggmap} -------------------------------
-### NOTE: this step can cost money $$$ (check Google pricing)
-
-# Register API key
-register_google(key = read_file("./britain-bombing-ww2/ref/keys/google-geocoding-api-key.txt"))
-
-# Look up geographic coordinates
-geocodes <- locations %>%
-  mutate(coord_data = map(lookup_location, geocode)) %>%
-  select(-lookup_location) %>%
-  unnest(coord_data)
-
-
-### Write to CSV ----------------------------------------------------------------------------------
-write_csv(geocodes, "./britain-bombing-ww2/ref/geocodes.csv")

### 2024-11-27 britain-bombing-ww2/src/clean-bombing-data.R

@@ -1,67 +1,0 @@
-
-### Load packages ---------------------------------------------------------------------------------
-library(tidyverse)
-library(readxl)
-library(janitor)
-
-
-### Load data and geocodes ------------------------------------------------------------------------
-bombings_raw <- read_xlsx("./britain-bombing-ww2/raw/Bombing-Britain-data.xlsx",
-                          guess_max = 1e5,
-                          na = c("", "N/A"))
-
-geocodes <- read_csv("./britain-bombing-ww2/ref/geocodes.csv")
-
-
-### Clean data ------------------------------------------------------------------------------------
-bombings_clean <- bombings_raw %>%
-  clean_names() %>%
-  # Add geographic coordinates (latitude and longitude)
-  left_join(geocodes, by = c("location", "country")) %>%
-  # Convert Excel dates to R date format
-  mutate(across(start_date:end_date, ~ excel_numeric_to_date(parse_number(.)))) %>%
-  # Parse numbers from casualty fields (they sometimes have values like "a few" or "unspecified")
-  mutate(across(killed:total_casualties, parse_number)) %>%
-  # Convert coordinates in the United States (clearly wrong) into missing values
-  mutate(lon = ifelse(lon < -25, NA, lon),
-         lat = ifelse(lon < -25, NA, lat)) %>%
-  # Add id field
-  mutate(bombing_id = row_number()) %>%
-  # Get rid of unnecessary fields
-  select(-volume_reference, -intelligence_summary_number, -link_to_page) %>%
-  # Re-order fields to be easier to work with
-  select(bombing_id, civil_defence_region, country, location,
-         lon, lat, start_date:time, killed:additional_notes)
-
-
-### Identify duplicate casualty records -----------------------------------------------------------
-# Identify duplicated casualty figures that are part of aggregations
-# We will create "casualty groups" that tick up by one when we conclude that the casualty data is not duplicated
-# This way, duplicated casualty figures will have the same "casualty group" ID, which we can use to prevent double-counting
-bombings_geocoded <- bombings_clean %>%
-  # The first row where one of these fields is different means a new casualty group
-  group_by(civil_defence_region, start_date, end_date, time, killed, injured, total_casualties) %>%
-  mutate(first_row = row_number() == 1) %>%
-  ungroup() %>%
-  # New casualty group when additional_notes fields doesn't include "Total" or "Entire", which are markers of aggregation
[…]

### 2024-11-28 scripts/01-test_simulated_data.R
add date, purpose, author

@@ -1,8 +1,8 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
+# Purpose: Tests the structure and validity of the datasets about the German 
+# bombing of England during the Second World War. 
+# Author: Dennis Netchitailo
+# Date: 27 November 2024
 # Contact: rohan.alexander@utoronto.ca
 # License: MIT
 # Pre-requisites: 

### 2024-11-28 scripts/02-download_data.R
added accidentally deleted raw file added process to download the datset, manually and automatically

@@ -1,21 +1,58 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from tacookson's GitHub repository.
+# Author: Dennis Netchitailo
+# Date: 27 November 2024
+# Contact: dennis.netchitailo@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
+# Pre-requisites: Download the "usethis" and "readr" libraries
 # Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
-library(opendatatoronto)
+
+#install.packages("usethis")
+#install.packages("readr")
+
+library(usethis)
 library(tidyverse)
-# [...UPDATE THIS...]
+library(readr)
+
+#install.packages("usethis")
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
 
+## ___ *Method #1 : Manual* ## ___
+
+# Go to the GitHub repository and manually download the .zip file 
+# Use the following link: https://github.com/tacookson/data
+# Location and open the .zip file
+
+# (1) Locate and move the "bombings.csv" file to the raw_data folder in the 
+# project following the path: data/01-raw_data
+
+# (2) Locate and move the "casualties.csv" file to the raw_data folder in the 
+# project following the path: data/01-raw_data
+
+## ___ Method #2 : Automatic* ## ___
+
+# Choose a directory on your computer to be the location of the .zip file
+# Insert the path of the directory where you want to save the .zip file
+#source_folder = ""
+source_folder = "C:\\Users\\Dennis Netchitailo\\Downloads"
+
+
+# Download the .zip file
+
+usethis::use_course(
+  url = 'https://github.com/tacookson/data/archive/refs/heads/master.zip',
+  destdir = destination)
+0
+source_csv_bombings <- file.path(source_folder, "bombings.csv")
+source_csv_casualties <- file.path(source_folder, "casualties.csv")
+
+#target_csv <- "data\\01-raw_data\\raw_data.csv"
[…]

### 2024-11-29 scripts/00-simulate_data.R
amended preamble. added data simulation (Poisson distribution)

@@ -1,9 +1,8 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulates 
+# Author: Dennis Netchitailo
+# Date: 28 November 2024
+# Contact: dennis.netchitailo@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
 # Any other information needed? Make sure you are in the `starter_folder` rproj
@@ -11,21 +10,38 @@
 
 #### Workspace setup ####
 library(tidyverse)
-set.seed(853)
-
+set.seed(1939)
 
 #### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
-)
+total_incidents = 32514
+num_casualties = 66000 + 220000 # ~ Dead + ~ Injured
+mean_casualties = num_casualties / total_incidents
+
+#Poisson Distribution Simulation
+simulated_casualties <- rpois(n = total_incidents, lambda = mean_casualties)
+
+# Verify Results
+total_simulated_casualties <- sum(simulated_casualties)
+mean_simulated_casualties <- mean(simulated_casualties)
+sd_simulated_casualties <- sd(simulated_casualties)
+
+# Output results
+cat("Total Simulated Casualties:", total_simulated_casualties, "\n")
+cat("Mean Simulated Casualties per Incident:", mean_simulated_casualties, "\n")
+cat("Standard Deviation of Simulated Casualties:", sd_simulated_casualties, "\n")
+
+
+
+
+
+
+
+
+
+
+
+
+
 
 # Political parties
 parties <- c("Labor", "Liberal", "Greens", "National", "Other")

### 2024-11-29 scripts/00-simulate_data.R
removed unnecessary code from example code

@@ -1,11 +1,10 @@
 #### Preamble ####
-# Purpose: Simulates 
+# Purpose: Simulates the distribution of casualties in the casualties data
 # Author: Dennis Netchitailo
 # Date: 28 November 2024
 # Contact: dennis.netchitailo@mail.utoronto.ca
-# License: MIT
+# License: --
 # Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
 
 
 #### Workspace setup ####
@@ -29,39 +28,6 @@ sd_simulated_casualties <- sd(simulated_casualties)
 cat("Total Simulated Casualties:", total_simulated_casualties, "\n")
 cat("Mean Simulated Casualties per Incident:", mean_simulated_casualties, "\n")
 cat("Standard Deviation of Simulated Casualties:", sd_simulated_casualties, "\n")
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-# Political parties
-parties <- c("Labor", "Liberal", "Greens", "National", "Other")
-
-# Create a dataset by randomly assigning states and parties to divisions
-analysis_data <- tibble(
-  division = paste("Division", 1:151),  # Add "Division" to make it a character
-  state = sample(
-    states,
-    size = 151,
-    replace = TRUE,
-    prob = c(0.25, 0.25, 0.15, 0.1, 0.1, 0.1, 0.025, 0.025) # Rough state population distribution
-  ),
-  party = sample(
-    parties,
-    size = 151,
-    replace = TRUE,
-    prob = c(0.40, 0.40, 0.05, 0.1, 0.05) # Rough party distribution
-  )
-)
 
 
 #### Save data ####

### 2024-11-29 scripts/01-test_simulated_data.R
updated preamble, added tests

@@ -3,7 +3,7 @@
 # bombing of England during the Second World War. 
 # Author: Dennis Netchitailo
 # Date: 27 November 2024
-# Contact: rohan.alexander@utoronto.ca
+# Contact: dennis.netchitailo@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
   # - The `tidyverse` package must be installed and loaded
@@ -26,46 +26,37 @@ if (exists("analysis_data")) {
 
 #### Test data ####
 
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
+
+
+# Check if the dataset has 32514 rows
+if (nrow(analysis_data) == 32514) {
+  message("Test Passed: The dataset has 32514 rows.")
 } else {
-  stop("Test Failed: The dataset does not have 151 rows.")
+  stop("Test Failed: The dataset does not have 32514 rows.")
 }
 
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
+# Check if the dataset has 4 columns
+if (ncol(analysis_data) == 4) {
+  message("Test Passed: The dataset has 4 columns.")
 } else {
-  stop("Test Failed: The dataset does not have 3 columns.")
+  stop("Test Failed: The dataset does not have 4 columns.")
 }
 
-# Check if all values in the 'division' column are unique
-if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
-  message("Test Passed: All values in 'division' are unique.")
+# Check if the 'location' column contains only valid locations of England's cities
+valid_cities <- c("London", "Manchester", "Liverpool", "Birmingham", "Coventry")
+
+if (all(analysis_data$location %in% valid_cities)) {
+  message("Test Passed: The 'location' column contains only valid English city names.")
 } else {
-  stop("Test Failed: The 'division' column contains duplicate values.")
+  stop("Test Failed: The 'location' column contains invalid city names.")
 }
 
-# Check if the 'state' column contains only valid Australian state names
-valid_states <- c("New South Wales", "Victoria", "Queensland", "South Australia", 
-                  "Western Australia", "Tasmania", "Northern Territory", 
-                  "Australian Capital Territory")
+# Check no negative casualties figures
+all(analysis_data$casualties >= 0) == TRUE
 
-if (all(analysis_data$state %in% valid_states)) {
[…]

### 2024-11-29 scripts/01-test_simulated_data.R
updated existing tests, added tests on date values, IDs, casualty outliers

@@ -65,11 +65,11 @@ if (all(!is.na(analysis_data))) {
   stop("Test Failed: The dataset contains missing values.")
 }
 
-# Check if there are no empty strings in 'incident_id', 'date', and 'location', 
+# Check if there are no empty strings in 'incident_id', 'location', 
 # and 'casualties' columns
-if (all(analysis_data$incident_id != "" & analysis_data$date != "" & analysis_data$location != "" &
+if (all(analysis_data$incident_id != "" & analysis_data$location != "" &
         analysis_data$casualties != "")) {
-  message("Test Passed: There are no empty strings in 'incident_id', 'date', 
+  message("Test Passed: There are no empty strings in 'incident_id', 
           'location' or 'casualties' .")
 } else {
   stop("Test Failed: There are empty strings in one or more columns.")
@@ -86,23 +86,33 @@ if (all(analysis_data$date >= start_date & analysis_data$date <= end_date)) {
   stop("Test Failed: Some dates fall outside the expected range.")
 }
 
-
-
-
-
-
-
-
 # Check if the 'party' column has at least two unique values
-if (n_distinct(analysis_data$party) >= 2) {
-  message("Test Passed: The 'party' column contains at least two unique values.")
+if (n_distinct(analysis_data$location) >= 2) {
+  message("Test Passed: The 'location' column contains at least two unique values.")
 } else {
-  stop("Test Failed: The 'party' column contains less than two unique values.")
-  
-  
-  
-
-  
-  
-  
+  stop("Test Failed: The 'location' column contains less than two unique values.")
 }
+
+# Check for IQR outliers
+outliers <- analysis_data %>%
+  filter(casualties > 17) 
+if (nrow(outliers) == 0) {
+  message("Test Passed: No extreme outliers in 'casualties'.")
+} else {
+  warning("Outliers detected in 'casualties':", nrow(outliers))
+}
+
+# Check all incident IDs are unique
+if (n_distinct(analysis_data$incident_id) == nrow(analysis_data)) {
+  message("Test Passed: All 'incident_id' values are unique.")
+} else {
+  stop("Test Failed: Duplicate 'incident_id' values found.")
+}
+
+# Check no missing date values
+if (all(!is.na(analysis_data$date))) {
+  message("Test Passed: No missing dates in the dataset.")
+} else {
+  stop("Test Failed: Missing dates detected in the dataset.")
+}
+

### 2024-11-29 scripts/02-download_data.R
updated preamble, removed unnecessary package installation code

@@ -3,21 +3,16 @@
 # Author: Dennis Netchitailo
 # Date: 27 November 2024
 # Contact: dennis.netchitailo@mail.utoronto.ca
-# License: MIT
+# License: --
 # Pre-requisites: Download the "usethis" and "readr" libraries
 # Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
 
-#install.packages("usethis")
-#install.packages("readr")
-
 library(usethis)
 library(tidyverse)
 library(readr)
-
-#install.packages("usethis")
 
 #### Download data ####
 
@@ -36,7 +31,7 @@ library(readr)
 ## ___ Method #2 : Automatic* ## ___
 
 # Choose a directory on your computer to be the location of the .zip file
-# Insert the path of the directory where you want to save the .zip file
+# Insert the pat 1945h of the directory where you want to save the .zip file
 #source_folder = ""
 source_folder = "C:\\Users\\Dennis Netchitailo\\Downloads"
 
@@ -46,7 +41,7 @@ source_folder = "C:\\Users\\Dennis Netchitailo\\Downloads"
 usethis::use_course(
   url = 'https://github.com/tacookson/data/archive/refs/heads/master.zip',
   destdir = destination)
-0
+
 source_csv_bombings <- file.path(source_folder, "bombings.csv")
 source_csv_casualties <- file.path(source_folder, "casualties.csv")
 

### 2024-11-29 scripts/02-download_data.R
improved paths, corrected directory location, added csv reading, removed unnecessary code

@@ -31,28 +31,25 @@ library(readr)
 ## ___ Method #2 : Automatic* ## ___
 
 # Choose a directory on your computer to be the location of the .zip file
-# Insert the pat 1945h of the directory where you want to save the .zip file
+# Insert the path of the directory where you want to save the .zip file
 #source_folder = ""
-source_folder = "C:\\Users\\Dennis Netchitailo\\Downloads"
+source_folder = "C:/Users/Dennis Netchitailo/Downloads/data-master/britain-bombing-ww2"
 
 
 # Download the .zip file
 
 usethis::use_course(
   url = 'https://github.com/tacookson/data/archive/refs/heads/master.zip',
-  destdir = destination)
+  destdir = source_folder)
 
 source_csv_bombings <- file.path(source_folder, "bombings.csv")
-source_csv_casualties <- file.path(source_folder, "casualties.csv")
+target_csv_bombings <- read.csv(source_csv_bombings) # Read the CSV file into R
 
-#target_csv <- "data\\01-raw_data\\raw_data.csv"
-target_csv_bombings <- "C:\Users\Dennis Netchitailo\Documents\STA304P3\data\01-raw_data\raw_data.csv"
-target_csv_casualties <- "C:\Users\Dennis Netchitailo\Documents\STA304P3\data\01-raw_data\raw_data.csv"
+source_csv_casualties <- file.path(source_folder, "casualties.csv")
+target_csv_casualties <- read.csv(source_csv_casualties) # Read the CSV file into R
 
 
 #### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
-
+write_csv(target_csv_bombings, "data/01-raw_data/bombings_data.csv") 
+write_csv(target_csv_casualties, "data/01-raw_data/casualties_data.csv")
          

### 2024-11-29 scripts/02-download_data.R
updated preamble pre-requisites

@@ -4,8 +4,7 @@
 # Date: 27 November 2024
 # Contact: dennis.netchitailo@mail.utoronto.ca
 # License: --
-# Pre-requisites: Download the "usethis" and "readr" libraries
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Download the `usethis`, `tidyverse`, and `readr` libraries
 
 
 #### Workspace setup ####

### 2024-11-29 scripts/00-simulate_data.R
updated pre-requisites, updated comments, removed unneeded code

@@ -4,31 +4,55 @@
 # Date: 28 November 2024
 # Contact: dennis.netchitailo@mail.utoronto.ca
 # License: --
-# Pre-requisites: The `tidyverse` package must be installed
-
+# Pre-requisites: The `tidyverse` and `here` packages must be installed
 
 #### Workspace setup ####
 library(tidyverse)
-set.seed(1939)
+library(here)
+set.seed(1939) # For reproducibility keep it as "1939"
 
 #### Simulate data ####
+
 total_incidents = 32514
 num_casualties = 66000 + 220000 # ~ Dead + ~ Injured
 mean_casualties = num_casualties / total_incidents
 
-#Poisson Distribution Simulation
-simulated_casualties <- rpois(n = total_incidents, lambda = mean_casualties)
+# Generate Dataset
+
+simulated_bombing_data <- tibble(
+  # Unique ID for each incident
+  incident_id = 1:total_incidents,
+  
+  # Randomly assign dates within a specific range
+  date = sample(seq.Date(from = as.Date("1939-09-01"), 
+                         to = as.Date("1940-05-10"), by = "day"), 
+                size = total_incidents, replace = TRUE),
+  
+  # Randomly assign locations (example locations)
+  location = sample(c("London", "Manchester", "Liverpool", "Birmingham", "Coventry"), 
+                    size = total_incidents, replace = TRUE),
+  
+  # Simulate casualties using a Poisson distribution
+  casualties = rpois(n = total_incidents, lambda = mean_casualties)
+)
+
+#### Save dataset ####
+output_path <- here("data/00-simulated_data/simulated_data.csv")
+write_csv(simulated_bombing_data, output_path)
+
+#### Verify dataset ####
+# Check structure
+print(simulated_bombing_data)
+cat("File saved at:", output_path, "\n")
+
+# Summary of casualties
+summary(simulated_bombing_data$casualties)
+
+#_______________________________________________________________________________
 
 # Verify Results
-total_simulated_casualties <- sum(simulated_casualties)
-mean_simulated_casualties <- mean(simulated_casualties)
-sd_simulated_casualties <- sd(simulated_casualties)
-
-# Output results
-cat("Total Simulated Casualties:", total_simulated_casualties, "\n")
-cat("Mean Simulated Casualties per Incident:", mean_simulated_casualties, "\n")
-cat("Standard Deviation of Simulated Casualties:", sd_simulated_casualties, "\n")
[…]

### 2024-11-29 scripts/00-simulate_data.R
removed unneeded code

@@ -48,11 +48,5 @@ cat("File saved at:", output_path, "\n")
 # Summary of casualties
 summary(simulated_bombing_data$casualties)
 
-#_______________________________________________________________________________
-
-# Verify Results
-#total_simulated_casualties <- sum(simulated_casualties)
-#mean_simulated_casualties <- mean(simulated_casualties)
-#sd_simulated_casualties <- sd(simulated_casualties)
 
 

### 2024-11-29 scripts/01-test_simulated_data.R
updated preamble, removed unnecessary spacing

@@ -4,7 +4,7 @@
 # Author: Dennis Netchitailo
 # Date: 27 November 2024
 # Contact: dennis.netchitailo@mail.utoronto.ca
-# License: MIT
+# License: --
 # Pre-requisites: 
   # - The `tidyverse` package must be installed and loaded
   # - 00-simulate_data.R must have been run
@@ -25,8 +25,6 @@ if (exists("analysis_data")) {
 
 
 #### Test data ####
-
-
 
 # Check if the dataset has 32514 rows
 if (nrow(analysis_data) == 32514) {

### 2024-11-29 scripts/01-test_simulated_data.R
improved formatting to make it clearer where new tests are being conducted removed duplicate test and unneeded code line

@@ -26,21 +26,21 @@ if (exists("analysis_data")) {
 
 #### Test data ####
 
-# Check if the dataset has 32514 rows
+## Check if the dataset has 32514 rows ##
 if (nrow(analysis_data) == 32514) {
   message("Test Passed: The dataset has 32514 rows.")
 } else {
   stop("Test Failed: The dataset does not have 32514 rows.")
 }
 
-# Check if the dataset has 4 columns
+## Check if the dataset has 4 columns ##
 if (ncol(analysis_data) == 4) {
   message("Test Passed: The dataset has 4 columns.")
 } else {
   stop("Test Failed: The dataset does not have 4 columns.")
 }
 
-# Check if the 'location' column contains only valid locations of England's cities
+## Check if the 'location' column contains only valid locations of England's cities ##
 valid_cities <- c("London", "Manchester", "Liverpool", "Birmingham", "Coventry")
 
 if (all(analysis_data$location %in% valid_cities)) {
@@ -49,14 +49,10 @@ if (all(analysis_data$location %in% valid_cities)) {
   stop("Test Failed: The 'location' column contains invalid city names.")
 }
 
-# Check no negative casualties figures
+## Check no negative casualties figures ##
 all(analysis_data$casualties >= 0) == TRUE
 
-# Check date range
-all(simulated_bombing_data$date >= as.Date("1939-09-01") & 
-      simulated_bombing_data$date <= as.Date("1940-05-10")) == TRUE
-
-# Check if there are any missing values in the dataset
+## Check if there are any missing values in the dataset ##
 if (all(!is.na(analysis_data))) {
   message("Test Passed: The dataset contains no missing values.")
 } else {
@@ -72,8 +68,8 @@ if (all(analysis_data$incident_id != "" & analysis_data$location != "" &
 } else {
   stop("Test Failed: There are empty strings in one or more columns.")
 }
-#analysis_data$incident_id != "" & analysis_data$date != "" &
 
+## Check all dates are within the set range ##
 # Define expected date range
 start_date <- as.Date("1939-09-01")
 end_date <- as.Date("1940-05-10")
@@ -91,7 +87,7 @@ if (n_distinct(analysis_data$location) >= 2) {
   stop("Test Failed: The 'location' column contains less than two unique values.")
 }
 
-# Check for IQR outliers
+## Check for IQR outliers ##
 outliers <- analysis_data %>%
   filter(casualties > 17) 
[…]

### 2024-11-29 scripts/02-download_data.R
made code work for non-local path added helper comment

@@ -31,8 +31,9 @@ library(readr)
 
 # Choose a directory on your computer to be the location of the .zip file
 # Insert the path of the directory where you want to save the .zip file
-#source_folder = ""
-source_folder = "C:/Users/Dennis Netchitailo/Downloads/data-master/britain-bombing-ww2"
+#source_folder = "C:/Users/Dennis Netchitailo/Downloads/data-master/britain-bombing-ww2"
+source_folder = "C:/Users/Dennis Netchitailo/Downloads"
+combined_path <- file.path(source_folder, "data-master/britain-bombing-ww2")
 
 
 # Download the .zip file
@@ -41,10 +42,10 @@ usethis::use_course(
   url = 'https://github.com/tacookson/data/archive/refs/heads/master.zip',
   destdir = source_folder)
 
-source_csv_bombings <- file.path(source_folder, "bombings.csv")
+source_csv_bombings <- file.path(combined_path, "bombings.csv")
 target_csv_bombings <- read.csv(source_csv_bombings) # Read the CSV file into R
 
-source_csv_casualties <- file.path(source_folder, "casualties.csv")
+source_csv_casualties <- file.path(combined_path, "casualties.csv")
 target_csv_casualties <- read.csv(source_csv_casualties) # Read the CSV file into R
 
 

### 2024-11-29 scripts/02-download_data.R
rewrote code to work for non-local paths added helper comment

@@ -31,8 +31,7 @@ library(readr)
 
 # Choose a directory on your computer to be the location of the .zip file
 # Insert the path of the directory where you want to save the .zip file
-#source_folder = "C:/Users/Dennis Netchitailo/Downloads/data-master/britain-bombing-ww2"
-source_folder = "C:/Users/Dennis Netchitailo/Downloads"
+source_folder = "" #Insert path here
 combined_path <- file.path(source_folder, "data-master/britain-bombing-ww2")
 
 

### 2024-11-29 scripts/02-download_data.R
added explanation on how to manually move and rename the datasets

@@ -23,9 +23,13 @@ library(readr)
 
 # (1) Locate and move the "bombings.csv" file to the raw_data folder in the 
 # project following the path: data/01-raw_data
+# Delete the "bombings_data.csv" file, and rename the "bombings.csv" file to
+# "bombings_data.csv"
 
 # (2) Locate and move the "casualties.csv" file to the raw_data folder in the 
 # project following the path: data/01-raw_data
+# Delete the "casualties_data.csv" file, and rename the "casualties.csv" file to
+# "casualties_data.csv"
 
 ## ___ Method #2 : Automatic* ## ___
 

### 2024-11-29 scripts/02-download_data.R
reordered the methods to download data: first automatic, then manual (the alternative method)

@@ -15,23 +15,7 @@ library(readr)
 
 #### Download data ####
 
-## ___ *Method #1 : Manual* ## ___
-
-# Go to the GitHub repository and manually download the .zip file 
-# Use the following link: https://github.com/tacookson/data
-# Location and open the .zip file
-
-# (1) Locate and move the "bombings.csv" file to the raw_data folder in the 
-# project following the path: data/01-raw_data
-# Delete the "bombings_data.csv" file, and rename the "bombings.csv" file to
-# "bombings_data.csv"
-
-# (2) Locate and move the "casualties.csv" file to the raw_data folder in the 
-# project following the path: data/01-raw_data
-# Delete the "casualties_data.csv" file, and rename the "casualties.csv" file to
-# "casualties_data.csv"
-
-## ___ Method #2 : Automatic* ## ___
+## ___ Method #1 : Automatic* ## ___
 
 # Choose a directory on your computer to be the location of the .zip file
 # Insert the path of the directory where you want to save the .zip file
@@ -55,4 +39,19 @@ target_csv_casualties <- read.csv(source_csv_casualties) # Read the CSV file int
 #### Save data ####
 write_csv(target_csv_bombings, "data/01-raw_data/bombings_data.csv") 
 write_csv(target_csv_casualties, "data/01-raw_data/casualties_data.csv")
-         
+
+## ___ *Method #2 : Manual* ## ___
+
+# Go to the GitHub repository and manually download the .zip file 
+# Use the following link: https://github.com/tacookson/data
+# Location and open the .zip file
+
+# (1) Locate and move the "bombings.csv" file to the raw_data folder in the 
+# project following the path: data/01-raw_data
+# Delete the "bombings_data.csv" file, and rename the "bombings.csv" file to
+# "bombings_data.csv"
+
+# (2) Locate and move the "casualties.csv" file to the raw_data folder in the 
+# project following the path: data/01-raw_data
+# Delete the "casualties_data.csv" file, and rename the "casualties.csv" file to
+# "casualties_data.csv"

### 2024-11-29 scripts/02-download_data.R
corrected spelling error

@@ -44,7 +44,7 @@ write_csv(target_csv_casualties, "data/01-raw_data/casualties_data.csv")
 
 # Go to the GitHub repository and manually download the .zip file 
 # Use the following link: https://github.com/tacookson/data
-# Location and open the .zip file
+# Locate and open the .zip file
 
 # (1) Locate and move the "bombings.csv" file to the raw_data folder in the 
 # project following the path: data/01-raw_data

### 2024-11-29 scripts/02-download_data.R
updated preamble

@@ -4,7 +4,10 @@
 # Date: 27 November 2024
 # Contact: dennis.netchitailo@mail.utoronto.ca
 # License: --
-# Pre-requisites: Download the `usethis`, `tidyverse`, and `readr` libraries
+# Pre-requisites: 
+  # - The `usethis` package must be installed and loaded
+  # - The `tidyverse` package must be installed and loaded
+  # - The `readr` package must be installed and loaded
 
 
 #### Workspace setup ####

### 2024-11-29 scripts/03-clean_data.R
updated preamble

@@ -1,10 +1,13 @@
 #### Preamble ####
 # Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
+# Author: Dennis Nethitailo
+# Date: November 29, 2024
+# Contact: dennis.netchitailo@mail.utoronto.ca 
+# License: --
+# Pre-requisites: 
+  # - 02-download_data.R must have been run
+  # - The `tidyverse` package must be installed and loaded
+  # - The `readr` package must be installed and loaded
 # Any other information needed? [...UPDATE THIS...]
 
 #### Workspace setup ####

### 2024-11-29 scripts/03-clean_data.R
updated package pre-req added library load

@@ -7,12 +7,12 @@
 # Pre-requisites: 
   # - 02-download_data.R must have been run
   # - The `tidyverse` package must be installed and loaded
-  # - The `readr` package must be installed and loaded
+  # - The `arrow` package must be installed and loaded
 # Any other information needed? [...UPDATE THIS...]
 
 #### Workspace setup ####
 library(tidyverse)
-
+library(arrow)
 #### Clean data ####
 raw_data <- read_csv("inputs/data/plane_data.csv")
 

### 2024-11-29 scripts/03-clean_data.R
updated preamble, added code to loan in data

@@ -1,6 +1,6 @@
 #### Preamble ####
 # Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Dennis Nethitailo
+# Author: Dennis Netchitailo
 # Date: November 29, 2024
 # Contact: dennis.netchitailo@mail.utoronto.ca 
 # License: --
@@ -14,7 +14,22 @@
 library(tidyverse)
 library(arrow)
 #### Clean data ####
+#source_folder = "" #Insert path here
+source_folder = "C:/Users/Dennis Netchitailo/Documents/STA304P3"
+combined_path <- file.path(source_folder, "data/01-raw_data/bombings_data.csv")
+
+
+
+csv_data <- read.csv("/data/01-raw_data/bombings_data.csv")
+
+
 raw_data <- read_csv("inputs/data/plane_data.csv")
+
+#write_parquet(data, "path/to/save/cleaned_data.parquet")
+
+
+
+
 
 cleaned_data <-
   raw_data |>

### 2024-11-30 scripts/03-clean_data.R
added load csv

@@ -12,24 +12,38 @@
 
 #### Workspace setup ####
 library(tidyverse)
-library(arrow)
+#library(arrow)
 #### Clean data ####
+
+# Copy and insert the path to the folder
 #source_folder = "" #Insert path here
 source_folder = "C:/Users/Dennis Netchitailo/Documents/STA304P3"
-combined_path <- file.path(source_folder, "data/01-raw_data/bombings_data.csv")
+#cobined_path <- file.path(source_folder, "data/01-raw_data/bombings.csv")
+
+load_csv <- function(file_name, base_folder = source_folder) {
+  file_path <- file.path(base_folder, file_name)
+  data <- read.csv(file_path)
+  return(data)
+}
+
+# Load bombings_data.csv
+bombings_data <- load_csv("data/01-raw_data/bombings_data.csv")
+
+# Load another CSV file, for example, casualties_data.csv
+casualties_data <- load_csv("data/01-raw_data/casualties_data.csv")
+
+# Inspect the datasets
+head(bombings_data)
+head(casualties_data)
 
 
-
-csv_data <- read.csv("/data/01-raw_data/bombings_data.csv")
-
-
-raw_data <- read_csv("inputs/data/plane_data.csv")
-
+#csv_data <- read.csv("/data/01-raw_data/bombings_data.csv")
+#raw_data <- read_csv("inputs/data/plane_data.csv")
 #write_parquet(data, "path/to/save/cleaned_data.parquet")
 
 
 
-
+# ______________________________________________________________________________
 
 cleaned_data <-
   raw_data |>

### 2024-11-30 scripts/03-clean_data.R
added code that creates a new column for proportion of killed / casualties

@@ -36,6 +36,22 @@ casualties_data <- load_csv("data/01-raw_data/casualties_data.csv")
 head(bombings_data)
 head(casualties_data)
 
+## Add new column ## 
+
+# Add a new column for the proportion of killed relative to total casualties
+casualties_data$proportion_killed <- casualties_data$killed / casualties_data$total_casualties
+
+# Handle cases where total_casualties is zero to avoid division by zero
+casualties_data$proportion_killed[casualties_data$total_casualties == 0] <- NA
+
+# Preview the updated dataset
+head(casualties_data)
+
+
+
+#### Save data ####
+write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+
 
 #csv_data <- read.csv("/data/01-raw_data/bombings_data.csv")
 #raw_data <- read_csv("inputs/data/plane_data.csv")
@@ -72,5 +88,4 @@ cleaned_data <-
          ) |> 
   tidyr::drop_na()
 
-#### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+

### 2024-11-30 scripts/03-clean_data.R
edited preamble, formatting, removed unneeded code

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
 # Author: Dennis Netchitailo
-# Date: November 29, 2024
+# Date: November 29 2024
 # Contact: dennis.netchitailo@mail.utoronto.ca 
 # License: --
 # Pre-requisites: 
@@ -26,36 +26,39 @@ load_csv <- function(file_name, base_folder = source_folder) {
   return(data)
 }
 
+#### Load Datasets ####
+
 # Load bombings_data.csv
 bombings_data <- load_csv("data/01-raw_data/bombings_data.csv")
 
-# Load another CSV file, for example, casualties_data.csv
-casualties_data <- load_csv("data/01-raw_data/casualties_data.csv")
+# Load casualties_data.csv
+cleaned_casualties_data <- load_csv("data/01-raw_data/casualties_data.csv")
 
-# Inspect the datasets
-head(bombings_data)
-head(casualties_data)
+### BOMBINGS Dataset ###
 
 ## Add new column ## 
 
 # Add a new column for the proportion of killed relative to total casualties
-casualties_data$proportion_killed <- casualties_data$killed / casualties_data$total_casualties
+cleaned_casualties_data$proportion_killed <- 
+  cleaned_casualties_data$killed / cleaned_casualties_data$total_casualties
 
 # Handle cases where total_casualties is zero to avoid division by zero
-casualties_data$proportion_killed[casualties_data$total_casualties == 0] <- NA
+cleaned_casualties_data$proportion_killed[cleaned_casualties_data$total_casualties == 0] <- NA
 
 # Preview the updated dataset
-head(casualties_data)
+head(cleaned_casualties_data)
 
+# Create dummy variable for Killed
+cleaned_casualties_data$has_killed <- ifelse(cleaned_casualties_data$killed > 0, 1, 0)
 
+# Create dummy variable for Injured
+cleaned_casualties_data$has_injured <- ifelse(cleaned_casualties_data$injured > 0, 1, 0)
+
+# Create dummy variable for Casualties
+cleaned_casualties_data$has_casualties <- ifelse(cleaned_casualties_data$total_casualties > 0, 1, 0)
 
 #### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")
-
-
-#csv_data <- read.csv("/data/01-raw_data/bombings_data.csv")
[…]

### 2024-11-30 scripts/03-clean_data.R
reoordered code cleaned bombings location var added count days for bombings date vars

@@ -18,7 +18,7 @@ library(tidyverse)
 # Copy and insert the path to the folder
 #source_folder = "" #Insert path here
 source_folder = "C:/Users/Dennis Netchitailo/Documents/STA304P3"
-#cobined_path <- file.path(source_folder, "data/01-raw_data/bombings.csv")
+#combined_path <- file.path(source_folder, "data/01-raw_data/bombings.csv")
 
 load_csv <- function(file_name, base_folder = source_folder) {
   file_path <- file.path(base_folder, file_name)
@@ -29,24 +29,21 @@ load_csv <- function(file_name, base_folder = source_folder) {
 #### Load Datasets ####
 
 # Load bombings_data.csv
-bombings_data <- load_csv("data/01-raw_data/bombings_data.csv")
+cleaned_bombings_data <- load_csv("data/01-raw_data/bombings_data.csv")
 
 # Load casualties_data.csv
 cleaned_casualties_data <- load_csv("data/01-raw_data/casualties_data.csv")
 
-### BOMBINGS Dataset ###
+#### 
+### CASUALTIES Dataset ###
 
-## Add new column ## 
+## Add new columns ## 
 
 # Add a new column for the proportion of killed relative to total casualties
 cleaned_casualties_data$proportion_killed <- 
   cleaned_casualties_data$killed / cleaned_casualties_data$total_casualties
-
 # Handle cases where total_casualties is zero to avoid division by zero
 cleaned_casualties_data$proportion_killed[cleaned_casualties_data$total_casualties == 0] <- NA
-
-# Preview the updated dataset
-head(cleaned_casualties_data)
 
 # Create dummy variable for Killed
 cleaned_casualties_data$has_killed <- ifelse(cleaned_casualties_data$killed > 0, 1, 0)
@@ -57,38 +54,21 @@ cleaned_casualties_data$has_injured <- ifelse(cleaned_casualties_data$injured > 
 # Create dummy variable for Casualties
 cleaned_casualties_data$has_casualties <- ifelse(cleaned_casualties_data$total_casualties > 0, 1, 0)
 
+### BOMBINGS Dataset ###
+
+# Remove whitespace before/after 
+cleaned_bombings_data$location <- trimws(cleaned_bombings_data$location)
+
+# Duration of bombing in days 
+cleaned_bombings_data$duration_days <- 
+  as.Date(cleaned_bombings_data$end_date) - as.Date(cleaned_bombings_data$start_date)
+
+
+
+
 #### Save data ####
[…]

### 2024-11-30 scripts/03-clean_data.R
for bombings dataset: added year column added month column added dummy for missing coordinates

@@ -63,6 +63,15 @@ cleaned_bombings_data$location <- trimws(cleaned_bombings_data$location)
 cleaned_bombings_data$duration_days <- 
   as.Date(cleaned_bombings_data$end_date) - as.Date(cleaned_bombings_data$start_date)
 
+# Year of occurrence
+cleaned_bombings_data$start_year <- format(as.Date(cleaned_bombings_data$start_date), "%Y")
+
+# Month of occurrence
+cleaned_bombings_data$start_month <- format(as.Date(cleaned_bombings_data$start_date), "%m")
+
+# Dummy for missing coordinates
+cleaned_bombings_data$has_missing_coords <- 
+  ifelse(is.na(cleaned_bombings_data$lon) | is.na(cleaned_bombings_data$lat), 1, 0)
 
 
 

### 2024-11-30 scripts/03-clean_data.R
removed unneeded code switched order of dataset loading
for bombings dataset,
added dummy time
removed notes column

@@ -7,18 +7,20 @@
 # Pre-requisites: 
   # - 02-download_data.R must have been run
   # - The `tidyverse` package must be installed and loaded
+  # - The `dplyr` package must be installed and loaded
   # - The `arrow` package must be installed and loaded
 # Any other information needed? [...UPDATE THIS...]
 
 #### Workspace setup ####
 library(tidyverse)
-#library(arrow)
+library(dplyr)
+library(arrow)
+
 #### Clean data ####
 
 # Copy and insert the path to the folder
 #source_folder = "" #Insert path here
 source_folder = "C:/Users/Dennis Netchitailo/Documents/STA304P3"
-#combined_path <- file.path(source_folder, "data/01-raw_data/bombings.csv")
 
 load_csv <- function(file_name, base_folder = source_folder) {
   file_path <- file.path(base_folder, file_name)
@@ -28,11 +30,12 @@ load_csv <- function(file_name, base_folder = source_folder) {
 
 #### Load Datasets ####
 
+# Load casualties_data.csv
+cleaned_casualties_data <- load_csv("data/01-raw_data/casualties_data.csv")
+
 # Load bombings_data.csv
 cleaned_bombings_data <- load_csv("data/01-raw_data/bombings_data.csv")
 
-# Load casualties_data.csv
-cleaned_casualties_data <- load_csv("data/01-raw_data/casualties_data.csv")
 
 #### 
 ### CASUALTIES Dataset ###
@@ -73,11 +76,27 @@ cleaned_bombings_data$start_month <- format(as.Date(cleaned_bombings_data$start_
 cleaned_bombings_data$has_missing_coords <- 
   ifelse(is.na(cleaned_bombings_data$lon) | is.na(cleaned_bombings_data$lat), 1, 0)
 
+# Dummy for unknown time
+cleaned_bombings_data$time_unknown <- ifelse(is.na(cleaned_bombings_data$time), 1, 0)
 
+# Drop additional notes column
+cleaned_bombings_data <- cleaned_bombings_data %>% select(-additional_notes)
 
 #### Save data ####
 write_csv(cleaned_casualties_data, "data/02-analysis_data/analysis_data_casualties.csv")
+write_csv(cleaned_bombings_data, "data/02-analysis_data/analysis_data_bombings.csv")
+
 
 ## Save as parquet ##
+# Load data
+cleaned_casualties_data <- read_csv("data/02-analysis_data/analysis_data_casualties.csv")
[…]

### 2024-11-30 scripts/03-clean_data.R
corrected day calculation removed unneeded code

@@ -64,7 +64,7 @@ cleaned_bombings_data$location <- trimws(cleaned_bombings_data$location)
 
 # Duration of bombing in days 
 cleaned_bombings_data$duration_days <- 
-  as.Date(cleaned_bombings_data$end_date) - as.Date(cleaned_bombings_data$start_date)
+  as.Date(cleaned_bombings_data$end_date) - as.Date(cleaned_bombings_data$start_date) + 1
 
 # Year of occurrence
 cleaned_bombings_data$start_year <- format(as.Date(cleaned_bombings_data$start_date), "%Y")
@@ -96,7 +96,5 @@ cleaned_bombings_data <- read_csv("data/02-analysis_data/analysis_data_bombings.
 write_parquet(cleaned_casualties_data, "data/02-analysis_data/analysis_data_casualties.parquet")
 write_parquet(cleaned_bombings_data, "data/02-analysis_data/analysis_data_bombings.parquet")
 
-head(casualties_data_parquet)
-head(bombings_data_parquet)
 
 

### 2024-11-30 scripts/04-test_analysis_data.R
updated preamble

@@ -1,10 +1,13 @@
 #### Preamble ####
 # Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
+# Author: Dennis Netchitailo
+# Date: 29 November 2024
+# Contact: dennis.netchitailo@mail.utoronto.ca 
+# License: --
 # Pre-requisites: [...UPDATE THIS...]
+  # - 03-clean_data.R must have been run
+  # - The `tidyverse` package must be installed and loaded
+  # - The `testthat` package must be installed and loaded
 # Any other information needed? [...UPDATE THIS...]
 
 

### 2024-11-30 scripts/03-clean_data.R
added convert bombing_id to int

@@ -59,6 +59,9 @@ cleaned_casualties_data$has_casualties <- ifelse(cleaned_casualties_data$total_c
 
 ### BOMBINGS Dataset ###
 
+# Make bombing_id integer
+cleaned_bombings_data$bombing_id <- as.integer(cleaned_bombings_data$bombing_id)
+
 # Remove whitespace before/after 
 cleaned_bombings_data$location <- trimws(cleaned_bombings_data$location)
 

### 2024-11-30 scripts/04-test_analysis_data.R
added tests

@@ -15,58 +15,89 @@
 library(tidyverse)
 library(testthat)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
+# Load data
+cleaned_data_bombings <- arrow::read_parquet(here::here("data/02-analysis_data/analysis_data_bombings.parquet"))
+cleaned_data_casualties <- arrow::read_parquet(here::here("data/02-analysis_data/analysis_data_casualties.parquet"))
 
+#data_bombings <- read_csv("data/02-analysis_data/analysis_data_bombings.csv")
+#data_casualties <- read_csv("data/02-analysis_data/analysis_data_casualties.csv")
 
 #### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
+
+# Ensure the data is loaded correctly
+#data_bombings <- read_csv("data/02-analysis_data/analysis_data_bombings.csv")
+#data_casualties <- read_csv("data/02-analysis_data/analysis_data_casualties.csv")
+
+# Convert necessary columns to the expected types (if not already)
+#data_bombings$start_date <- as.Date(data_bombings$start_date, format="%Y-%m-%d")
+#data_bombings$end_date <- as.Date(data_bombings$end_date, format="%Y-%m-%d")
+
+#### BOMBING DATASET TESTS ####
+
+test_that("bombing_id is of class integer", {
+  expect_type(data_bombings$bombing_id, "integer")
 })
 
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
+test_that("civil_defense_region is of class character", {
+  expect_type(data_bombings$civil_defense_region, "character")
 })
 
-# Test that the 'division' column is character type
-test_that("'division' is character", {
-  expect_type(analysis_data$division, "character")
+test_that("country is of class character", {
+  expect_type(data_bombings$country, "character")
 })
 
-# Test that the 'party' column is character type
-test_that("'party' is character", {
-  expect_type(analysis_data$party, "character")
+test_that("location is of class character", {
+  expect_type(data_bombings$location, "character")
 })
 
-# Test that the 'state' column is character type
-test_that("'state' is character", {
-  expect_type(analysis_data$state, "character")
+test_that("lon is of class numeric", {
+  expect_type(data_bombings$lon, "numeric")
 })
 
-# Test that there are no missing values in the dataset
[…]

### 2024-11-30 scripts/04-test_analysis_data.R
corrected test df corrected expected class types

@@ -14,90 +14,82 @@
 #### Workspace setup ####
 library(tidyverse)
 library(testthat)
+library(arrow)
+library(here)
 
 # Load data
 cleaned_data_bombings <- arrow::read_parquet(here::here("data/02-analysis_data/analysis_data_bombings.parquet"))
 cleaned_data_casualties <- arrow::read_parquet(here::here("data/02-analysis_data/analysis_data_casualties.parquet"))
 
-#data_bombings <- read_csv("data/02-analysis_data/analysis_data_bombings.csv")
-#data_casualties <- read_csv("data/02-analysis_data/analysis_data_casualties.csv")
-
 #### Test data ####
-
-# Ensure the data is loaded correctly
-#data_bombings <- read_csv("data/02-analysis_data/analysis_data_bombings.csv")
-#data_casualties <- read_csv("data/02-analysis_data/analysis_data_casualties.csv")
-
-# Convert necessary columns to the expected types (if not already)
-#data_bombings$start_date <- as.Date(data_bombings$start_date, format="%Y-%m-%d")
-#data_bombings$end_date <- as.Date(data_bombings$end_date, format="%Y-%m-%d")
 
 #### BOMBING DATASET TESTS ####
 
-test_that("bombing_id is of class integer", {
-  expect_type(data_bombings$bombing_id, "integer")
+test_that("bombing_id is of class double", {
+  expect_type(cleaned_data_bombings$bombing_id, "double")
 })
 
 test_that("civil_defense_region is of class character", {
-  expect_type(data_bombings$civil_defense_region, "character")
+  expect_type(cleaned_data_bombings$civil_defence_region, "character")
 })
 
 test_that("country is of class character", {
-  expect_type(data_bombings$country, "character")
+  expect_type(cleaned_data_bombings$country, "character")
 })
 
 test_that("location is of class character", {
-  expect_type(data_bombings$location, "character")
+  expect_type(cleaned_data_bombings$location, "character")
 })
 
-test_that("lon is of class numeric", {
-  expect_type(data_bombings$lon, "numeric")
+test_that("lon is of class double", {
+  expect_type(cleaned_data_bombings$lon, "double")
 })
 
-test_that("lat is of class numeric", {
-  expect_type(data_bombings$lat, "numeric")
+test_that("lat is of class double", {
+  expect_type(cleaned_data_bombings$lat, "double")
 })
 
 test_that("start_date is of class Date", {
-  expect_s3_class(data_bombings$start_date, "Date")
[…]

### 2024-11-30 scripts/03-clean_data.R
removed uneeded code

@@ -85,6 +85,26 @@ cleaned_bombings_data$time_unknown <- ifelse(is.na(cleaned_bombings_data$time), 
 # Drop additional notes column
 cleaned_bombings_data <- cleaned_bombings_data %>% select(-additional_notes)
 
+# _______________________
+
+# Convert bombing_id to integer
+cleaned_bombings_data$bombing_id <- as.integer(cleaned_bombings_data$bombing_id)
+
+# Convert casualty_group, duration_days, start_year to integer
+cleaned_bombings_data$casualty_group <- as.integer(cleaned_bombings_data$casualty_group)
+cleaned_bombings_data$duration_days <- as.integer(cleaned_bombings_data$duration_days)
+cleaned_bombings_data$start_year <- as.integer(cleaned_bombings_data$start_year)
+
+# Convert has_missing_coords, time_unknown to integer
+cleaned_bombings_data$has_missing_coords <- as.integer(cleaned_bombings_data$has_missing_coords)
+cleaned_bombings_data$time_unknown <- as.integer(cleaned_bombings_data$time_unknown)
+
+# For start_month, if you want it as an integer (month as 1-12), convert it
+cleaned_bombings_data$start_month <- as.integer(cleaned_bombings_data$start_month)
+
+
+#________________________
+
 #### Save data ####
 write_csv(cleaned_casualties_data, "data/02-analysis_data/analysis_data_casualties.csv")
 write_csv(cleaned_bombings_data, "data/02-analysis_data/analysis_data_bombings.csv")

### 2024-11-30 scripts/03-clean_data.R
removed unneeded code

@@ -85,26 +85,6 @@ cleaned_bombings_data$time_unknown <- ifelse(is.na(cleaned_bombings_data$time), 
 # Drop additional notes column
 cleaned_bombings_data <- cleaned_bombings_data %>% select(-additional_notes)
 
-# _______________________
-
-# Convert bombing_id to integer
-cleaned_bombings_data$bombing_id <- as.integer(cleaned_bombings_data$bombing_id)
-
-# Convert casualty_group, duration_days, start_year to integer
-cleaned_bombings_data$casualty_group <- as.integer(cleaned_bombings_data$casualty_group)
-cleaned_bombings_data$duration_days <- as.integer(cleaned_bombings_data$duration_days)
-cleaned_bombings_data$start_year <- as.integer(cleaned_bombings_data$start_year)
-
-# Convert has_missing_coords, time_unknown to integer
-cleaned_bombings_data$has_missing_coords <- as.integer(cleaned_bombings_data$has_missing_coords)
-cleaned_bombings_data$time_unknown <- as.integer(cleaned_bombings_data$time_unknown)
-
-# For start_month, if you want it as an integer (month as 1-12), convert it
-cleaned_bombings_data$start_month <- as.integer(cleaned_bombings_data$start_month)
-
-
-#________________________
-
 #### Save data ####
 write_csv(cleaned_casualties_data, "data/02-analysis_data/analysis_data_casualties.csv")
 write_csv(cleaned_bombings_data, "data/02-analysis_data/analysis_data_bombings.csv")

### 2024-11-30 scripts/04-test_analysis_data.R
added tests for casualties dataset

@@ -16,6 +16,8 @@ library(tidyverse)
 library(testthat)
 library(arrow)
 library(here)
+library(validate)
+library(pointblank)
 
 # Load data
 cleaned_data_bombings <- arrow::read_parquet(here::here("data/02-analysis_data/analysis_data_bombings.parquet"))
@@ -92,4 +94,21 @@ test_that("has_missing_coords is of class double", {
 
 test_that("time_unknown is of class double", {
   expect_type(cleaned_data_bombings$time_unknown, "double")
+})
+
+#### BOMBING DATASET TESTS ####
+
+# Values >0 for killed
+test_that("All values in 'killed' are greater than 0", {
+  expect_true(all(cleaned_data_casualties$killed >= 0, na.rm = TRUE))
+})
+
+# Values >0 for injured
+test_that("All values in 'injured' are greater than 0", {
+  expect_true(all(cleaned_data_casualties$injured >= 0, na.rm = TRUE))
+})
+
+# Values >0 for casualties
+test_that("All values in 'total_casualties' are greater than 0", {
+  expect_true(all(cleaned_data_casualties$total_casualties >= 0, na.rm = TRUE))
 })

### 2024-11-30 scripts/04-test_analysis_data.R
updated preamble removed unneeded load libraries

@@ -8,6 +8,8 @@
   # - 03-clean_data.R must have been run
   # - The `tidyverse` package must be installed and loaded
   # - The `testthat` package must be installed and loaded
+  # - The `arrow` package must be installed and loaded
+  # - The `here` package must be installed and loaded
 # Any other information needed? [...UPDATE THIS...]
 
 
@@ -16,8 +18,6 @@ library(tidyverse)
 library(testthat)
 library(arrow)
 library(here)
-library(validate)
-library(pointblank)
 
 # Load data
 cleaned_data_bombings <- arrow::read_parquet(here::here("data/02-analysis_data/analysis_data_bombings.parquet"))

### 2024-11-30 scripts/05-exploratory_data_analysis.R
updated preamble

@@ -1,10 +1,13 @@
 #### Preamble ####
 # Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
+# Author: Dennis Netchitailo
+# Date: 30 November 2024
+# Contact: dennis.netchitailo@mail.utoronto.ca 
+# License: --
 # Pre-requisites: [...UPDATE THIS...]
+  # - The `tidyverse` package must be installed and loaded
+  # - The `rstanarm` package must be installed and loaded
+
 # Any other information needed? [...UPDATE THIS...]
 
 

### 2024-11-30 scripts/06-model_data.R
updated preamble

@@ -1,10 +1,12 @@
 #### Preamble ####
 # Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
+# Author: Dennis Netchitailo
+# Date: 30 November 2024
+# Contact: dennis.netchitailo@mail.utoronto.ca 
+# License: --
 # Pre-requisites: [...UPDATE THIS...]
+  # - The `tidyverse` package must be installed and loaded
+  # - The `rstanarm` package must be installed and loaded
 # Any other information needed? [...UPDATE THIS...]
 
 

### 2024-11-30 scripts/05-exploratory_data_analysis.R
updated preamble

@@ -4,7 +4,8 @@
 # Date: 30 November 2024
 # Contact: dennis.netchitailo@mail.utoronto.ca 
 # License: --
-# Pre-requisites: [...UPDATE THIS...]
+# Pre-requisites: 
+  # - 03-clean_data.R must have been run
   # - The `tidyverse` package must be installed and loaded
   # - The `rstanarm` package must be installed and loaded
 

### 2024-11-30 scripts/05-exploratory_data_analysis.R
added summary stats for bombings & casualties datasets

@@ -16,10 +16,74 @@
 library(tidyverse)
 library(rstanarm)
 
+source_folder = "C:/Users/Dennis Netchitailo/Documents/STA304P3"
+
+load_csv <- function(file_name, base_folder = source_folder) {
+  file_path <- file.path(base_folder, file_name)
+  data <- read.csv(file_path)
+  return(data)
+}
+
+#### Load Datasets ####
+
+# Load casualties_data.csv
+casualties_analysis_data <- load_csv("data/02-analysis_data/analysis_data_casualties.csv")
+
+# Load bombings_data.csv
+bombings_analysis_data <- load_csv("data/02-analysis_data/analysis_data_bombings.csv")
+
+
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+#analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
 
 ### Model data ####
+
+## CASUALTIES DATASET ##
+
+# Sum Killed
+sum(casualties_analysis_data$killed, na.rm = TRUE)
+
+# Sum Injured
+sum(casualties_analysis_data$injured, na.rm = TRUE)
+
+# Sum Casualties
+sum(casualties_analysis_data$total_casualties, na.rm = TRUE)
+
+# Sum of difference between Casualties and Killed & Injured
+sum(casualties_analysis_data$total_casualties, na.rm = TRUE) - 
+  (sum(casualties_analysis_data$killed, na.rm = TRUE)
+ + sum(casualties_analysis_data$injured, na.rm = TRUE))
+
+
+## BOMBINGS DATASET ##
+
+# Count Countries
+bombings_analysis_data %>%
+  count(country)
+
+# Earliest and latest bombing dates
+range(bombings_analysis_data$start_date, na.rm = TRUE)
+
+# Bombings per year
+bombings_analysis_data %>%
+  count(start_year)
+
+# Bombings per month (seasonal analysis)
+bombings_analysis_data %>%
+  count(start_month)
+
+# Number of bombings by civil defense region
+bombings_analysis_data %>%
+  count(civil_defence_region)
+
+# Count bombings with missing coordinates
+sum(is.na(bombings_analysis_data$lon) | is.na(bombings_analysis_data$lat))
+
+summary(bombings_analysis_data$duration_days)
+
+
+# __________________________________________________________________
+
 first_model <-
   stan_glm(
     formula = flying_time ~ length + width,

### 2024-12-01 scripts/03-clean_data.R
merged datasets into combined_data

@@ -37,7 +37,6 @@ cleaned_casualties_data <- load_csv("data/01-raw_data/casualties_data.csv")
 cleaned_bombings_data <- load_csv("data/01-raw_data/bombings_data.csv")
 
 
-#### 
 ### CASUALTIES Dataset ###
 
 ## Add new columns ## 
@@ -89,11 +88,47 @@ cleaned_bombings_data <- cleaned_bombings_data %>% select(-additional_notes)
 write_csv(cleaned_casualties_data, "data/02-analysis_data/analysis_data_casualties.csv")
 write_csv(cleaned_bombings_data, "data/02-analysis_data/analysis_data_bombings.csv")
 
+##### COMBINED DATASET #####
+
+#### Load Datasets ####
+
+# Load casualties_data.csv
+casualties_data <- load_csv("data/02-analysis_data/analysis_data_casualties.csv")
+
+# Load bombings_data.csv
+bombings_data <- load_csv("data/02-analysis_data/analysis_data_bombings.csv")
+
+## Combine ## 
+
+aggregated_bombings <- bombings_data %>%
+  group_by(casualty_group) %>%
+  summarize(
+    total_incidents = n(),
+    civil_defense_region = first(civil_defence_region),
+    country = first(country),
+    start_date = first(start_date),
+    end_date = first(end_date),
+    time = first(time),
+    duration_days = first(duration_days),
+    start_year = first(start_year),
+    start_month = first(start_month),
+    time_unknown = first(time_unknown),
+    .groups = "drop"
+  )
+
+combined_data <- aggregated_bombings %>%
+  left_join(casualties_data, by = "casualty_group")
+
+
+## Write Data ##
+write_csv(combined_data, "data/02-analysis_data/combined_data.csv")
+message("File successfully written!")
 
 ## Save as parquet ##
 # Load data
 cleaned_casualties_data <- read_csv("data/02-analysis_data/analysis_data_casualties.csv")
 cleaned_bombings_data <- read_csv("data/02-analysis_data/analysis_data_bombings.csv")
+
 
 # Save as Parquet
 write_parquet(cleaned_casualties_data, "data/02-analysis_data/analysis_data_casualties.parquet")

### 2024-12-01 scripts/03-clean_data.R
updated comments

@@ -90,7 +90,7 @@ write_csv(cleaned_bombings_data, "data/02-analysis_data/analysis_data_bombings.c
 
 ##### COMBINED DATASET #####
 
-#### Load Datasets ####
+#### Load Dataset ####
 
 # Load casualties_data.csv
 casualties_data <- load_csv("data/02-analysis_data/analysis_data_casualties.csv")
@@ -120,7 +120,7 @@ combined_data <- aggregated_bombings %>%
   left_join(casualties_data, by = "casualty_group")
 
 
-## Write Data ##
+## Save Data ##
 write_csv(combined_data, "data/02-analysis_data/combined_data.csv")
 message("File successfully written!")
 

### 2024-12-01 scripts/03-clean_data.R
added save as parquet for combined_data

@@ -128,11 +128,13 @@ message("File successfully written!")
 # Load data
 cleaned_casualties_data <- read_csv("data/02-analysis_data/analysis_data_casualties.csv")
 cleaned_bombings_data <- read_csv("data/02-analysis_data/analysis_data_bombings.csv")
+combined_data <- read_csv("data/02-analysis_data/combined_data.csv")
 
 
 # Save as Parquet
 write_parquet(cleaned_casualties_data, "data/02-analysis_data/analysis_data_casualties.parquet")
 write_parquet(cleaned_bombings_data, "data/02-analysis_data/analysis_data_bombings.parquet")
+write_parquet(combined_data, "data/02-analysis_data/combined_data.parquet")
 
 
 

### 2024-12-02 scripts/03-clean_data.R
added rows cleaned rows with invalid Killed/Casualties proportions

@@ -119,6 +119,40 @@ aggregated_bombings <- bombings_data %>%
 combined_data <- aggregated_bombings %>%
   left_join(casualties_data, by = "casualty_group")
 
+# Convert casualty_group to integer type
+combined_data <- combined_data %>%
+  mutate(casualty_group = as.integer(casualty_group))
+
+
+combined_data <- combined_data %>%
+  mutate(
+    proportion_killed = ifelse(total_casualties > 0, killed / total_casualties, NA)
+  )
+
+# Remove rows with invalid proportions
+cleaned_data <- combined_data %>%
+  filter(is.na(proportion_killed) | (proportion_killed >= 0 & proportion_killed <= 1))
+combined_data <- cleaned_data
+
+# Dummy variable for Time
+cleaned_data <- combined_data %>%
+  mutate(
+    # Standardize the time column
+    time = case_when(
+      str_to_lower(time) == "day" ~ "Day",
+      str_to_lower(time) == "night" ~ "Night",
+      TRUE ~ "Unspecified"  # Handle unspecified or other unexpected values
+    ),
+    # Create time_binary: 0 = Day, 1 = Night, NA for Unspecified
+    time_binary = ifelse(time == "Night", 1,
+                         ifelse(time == "Day", 0, NA))
+  )
+
+
+# View the cleaned dataset
+head(cleaned_data)
+
+combined_data <- cleaned_data
 
 ## Save Data ##
 write_csv(combined_data, "data/02-analysis_data/combined_data.csv")

### 2024-12-02 scripts/04-test_analysis_data.R
corrected expected classes

@@ -20,95 +20,100 @@ library(arrow)
 library(here)
 
 # Load data
-cleaned_data_bombings <- arrow::read_parquet(here::here("data/02-analysis_data/analysis_data_bombings.parquet"))
-cleaned_data_casualties <- arrow::read_parquet(here::here("data/02-analysis_data/analysis_data_casualties.parquet"))
+combined_data <- arrow::read_parquet(here::here("data/02-analysis_data/combined_data.parquet"))
 
 #### Test data ####
 
-#### BOMBING DATASET TESTS ####
+#### COMBINED DATASET TESTS ####
 
-test_that("bombing_id is of class double", {
-  expect_type(cleaned_data_bombings$bombing_id, "double")
+# Test casualty-related columns
+test_that("All values in 'killed' are greater than or equal to 0", {
+  expect_true(all(combined_data$killed >= 0, na.rm = TRUE))
+})
+
+test_that("All values in 'injured' are greater than or equal to 0", {
+  expect_true(all(combined_data$injured >= 0, na.rm = TRUE))
+})
+
+test_that("All values in 'total_casualties' are greater than or equal to 0", {
+  expect_true(all(combined_data$total_casualties >= 0, na.rm = TRUE))
+})
+
+# Test bombing-related columns
+test_that("casualty_group is of class double", {
+  expect_type(combined_data$casualty_group, "double")
+})
+
+test_that("total_incidents is of class double", {
+  expect_type(combined_data$total_incidents, "double")
 })
 
 test_that("civil_defense_region is of class character", {
-  expect_type(cleaned_data_bombings$civil_defence_region, "character")
+  expect_type(combined_data$civil_defense_region, "character")
 })
 
 test_that("country is of class character", {
-  expect_type(cleaned_data_bombings$country, "character")
+  expect_type(combined_data$country, "character")
 })
 
-test_that("location is of class character", {
-  expect_type(cleaned_data_bombings$location, "character")
-})
-
-test_that("lon is of class double", {
-  expect_type(cleaned_data_bombings$lon, "double")
-})
-
-test_that("lat is of class double", {
-  expect_type(cleaned_data_bombings$lat, "double")
-})
-
-test_that("start_date is of class Date", {
-  expect_s3_class(cleaned_data_bombings$start_date, "Date")
+test_that("start_date is of class double", {
+  expect_type(combined_data$start_date, "double")
 })
 
[…]

### 2024-12-02 scripts/05-exploratory_data_analysis.R
added summary statistics

@@ -15,6 +15,7 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(dplyr)
 
 source_folder = "C:/Users/Dennis Netchitailo/Documents/STA304P3"
 
@@ -54,6 +55,35 @@ sum(casualties_analysis_data$total_casualties, na.rm = TRUE) -
   (sum(casualties_analysis_data$killed, na.rm = TRUE)
  + sum(casualties_analysis_data$injured, na.rm = TRUE))
 
+# Mean Killed
+mean(casualties_analysis_data$killed, na.rm = TRUE)
+
+# Mean Injured
+mean(casualties_analysis_data$injured, na.rm = TRUE)
+
+# Mean Casualties
+mean(casualties_analysis_data$total_casualties, na.rm = TRUE)
+
+# Distribution of casualties
+summary(casualties_analysis_data$total_casualties)
+
+# Distribution of killed across groups
+casualties_analysis_data %>%
+  summarize(
+    min_killed = min(killed, na.rm = TRUE),
+    max_killed = max(killed, na.rm = TRUE),
+    mean_killed = mean(killed, na.rm = TRUE),
+    median_killed = median(killed, na.rm = TRUE)
+  )
+
+# Distribution of injured across groups
+casualties_analysis_data %>%
+  summarize(
+    min_injured = min(injured, na.rm = TRUE),
+    max_injured = max(injured, na.rm = TRUE),
+    mean_injured = mean(injured, na.rm = TRUE),
+    median_injured = median(injured, na.rm = TRUE)
+  )
 
 ## BOMBINGS DATASET ##
 
@@ -80,6 +110,18 @@ bombings_analysis_data %>%
 sum(is.na(bombings_analysis_data$lon) | is.na(bombings_analysis_data$lat))
 
 summary(bombings_analysis_data$duration_days)
+
+## COMBINED Summary Statistics ##
+
+# Count incidents per casualty group
+incident_counts <- bombings_analysis_data %>%
+  group_by(casualty_group) %>%
+  summarize(total_incidents = n(), .groups = "drop")
+
+
+combined_data %>% 
+  summarize(avg_incidents_per_group = mean(total_incidents, na.rm = TRUE))
+
 
 
 # __________________________________________________________________

### 2024-12-02 scripts/05-exploratory_data_analysis.R
updated preamble

@@ -8,6 +8,7 @@
   # - 03-clean_data.R must have been run
   # - The `tidyverse` package must be installed and loaded
   # - The `rstanarm` package must be installed and loaded
+  # - The `ggplot2` package must be installed and loaded
 
 # Any other information needed? [...UPDATE THIS...]
 
@@ -16,6 +17,8 @@
 library(tidyverse)
 library(rstanarm)
 library(dplyr)
+library(ggplot2)
+
 
 source_folder = "C:/Users/Dennis Netchitailo/Documents/STA304P3"
 
@@ -126,6 +129,32 @@ combined_data %>%
 
 # __________________________________________________________________
 
+first_model <- stan_glm(
+  total_casualties ~ time_binary,
+  data = combined_data,
+  family = gaussian(),  # Change family if appropriate
+  prior = normal(0, 2),  # Prior for the slope
+  prior_intercept = normal(0, 5),  # Prior for the intercept
+  chains = 4,  # Number of chains
+  iter = 2000,  # Number of iterations
+  seed = 12345  # For reproducibility
+)
+
+first_model
+
+# Summarize the model
+summary(first_model)
+
+library(ggplot2)
+
+ggplot(combined_data, aes(x = as.factor(time_binary), y = total_casualties)) +
+  geom_boxplot() +
+  labs(x = "Time (0 = Day, 1 = Night)", y = "Total Casualties") +
+  theme_minimal()
+
+
+
+# __________________________________________________________________
 first_model <-
   stan_glm(
     formula = flying_time ~ length + width,

### 2024-12-02 scripts/05-exploratory_data_analysis.R
updated preamble

@@ -9,6 +9,7 @@
   # - The `tidyverse` package must be installed and loaded
   # - The `rstanarm` package must be installed and loaded
   # - The `ggplot2` package must be installed and loaded
+  # - The `brms` package must be installed and loaded
 
 # Any other information needed? [...UPDATE THIS...]
 
@@ -18,6 +19,7 @@ library(tidyverse)
 library(rstanarm)
 library(dplyr)
 library(ggplot2)
+library(brms)
 
 
 source_folder = "C:/Users/Dennis Netchitailo/Documents/STA304P3"
@@ -145,11 +147,60 @@ first_model
 # Summarize the model
 summary(first_model)
 
-library(ggplot2)
+# Create predicted values for the model
+combined_data_first_model <- combined_data
+combined_data_first_model$predicted <- posterior_predict(first_model, summary = TRUE)
 
+# Plot data and regression line
+ggplot(combined_data, aes(x = time_binary, y = total_casualties)) +
+  geom_point(alpha = 0.5) +  # Actual data points
+  geom_smooth(aes(y = predicted), method = "lm", color = "blue") +  # Regression line
+  labs(title = "Regression Model: Total Casualties vs Time",
+       x = "Time (0 = Day, 1 = Night)",
+       y = "Total Casualties") +
+  theme_minimal()
+
+
+
+
+# Boxplot of Data
 ggplot(combined_data, aes(x = as.factor(time_binary), y = total_casualties)) +
   geom_boxplot() +
   labs(x = "Time (0 = Day, 1 = Night)", y = "Total Casualties") +
+  theme_minimal()
+
+
+# Fit a Zero-Inflated Negative Binomial model
+zero_inflated_model <- brm(
+  formula = total_casualties ~ time_binary + (1 | casualty_group),  # Add random effects if needed
+  data = combined_data,
+  family = zero_inflated_negbinomial(link = "log"),  # Zero-inflated negative binomial
+  prior = c(
+    prior(normal(0, 5), class = "b"),  # Priors for fixed effects
+    prior(cauchy(0, 2), class = "sd")  # Priors for random effects if present
+  ),
+  chains = 4,
+  iter = 2000,
+  cores = 4,
+  seed = 123
+)
+
+# Summarize the zero-inflated model
+summary(zero_inflated_model)
+
+# Posterior predictive checks
+pp_check(zero_inflated_model)
+
+# Generate predicted values
[…]

### 2024-12-02 paper/paper.qmd
updated preable first version of abstract updated required libraries

@@
   - Dennis Netchitailo
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
-date: today
+thanks: "Code and data are available at: 
+[https://github.com/dennisnetchitailo/STA304P3]
+(https://github.com/tacookson/data/tree/master/britain-bombing-ww2)."
+date: 2 December 2024
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study models the German bombing campaign of Britain during the Second World War by analyzing the data collected by the Home Security War Room from 1939 to 1945 to summarise air raid damage.
+I use a multiple linear regression model to estimate the impact of the time an attack occurred on casualties."
 format: pdf
@@
 library(palmerpenguins)
+library(arrow)
+library(here)
+
 ```

### 2024-12-02 paper/paper.qmd
updated libraries
updated introduction

@@
 library(here)
+library(readr)
+library(usethis)
+library(dplyr)
+library(testthat)
+library(rstanarm)
+library(ggplot2)
+library(brms)
 
@@
 
-Overview paragraph
+////Overview paragraph////
+The Ministry of Home Security was the British government department established in 1939 to direct the response to German air raids.
+Part of its responsibility was tracking data related to air raids.
+This data
 
-Estimand paragraph
+////Estimand paragraph////
+The estimand is the number of casualties sustained during an attack.
+I use the Home Security data with a multiple linear regression model to estimate the difference in casualties during nighttime air raids compared to daytime air raids
 
-Results paragraph
+////Results paragraph////
 
-Why it matters paragraph
+////Why it matters paragraph////
 

### 2024-12-03 paper/paper.qmd
added intro edited data overview

@@
 thanks: "Code and data are available at: 
-[https://github.com/dennisnetchitailo/STA304P3]
-(https://github.com/tacookson/data/tree/master/britain-bombing-ww2)."
+[https://github.com/dennisnetchitailo/STA304P3](https://github.com/dennisnetchitailo/STA304P3)."
 date: 2 December 2024
@@
 ////Overview paragraph////
+On September 1st, 1939, the German armed forces crossed the Polish border. The United Kingdom declared war on Germany on September 3rd, 1939. 
 The Ministry of Home Security was the British government department established in 1939 to direct the response to German air raids.
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+I used the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
 

### 2024-12-03 paper/paper.qmd
updated abstract

@@
 date-format: long
-abstract: "This study models the German bombing campaign of Britain during the Second World War by analyzing the data collected by the Home Security War Room from 1939 to 1945 to summarise air raid damage.
-I use a multiple linear regression model to estimate the impact of the time an attack occurred on casualties."
+abstract: "With wars raging across the world, notably, between Russia and Ukraine, and aerial bombardment using aircraft or missiles a major cause of civilian casualties far from the frontlines, understanding, based on historical data, whether either nighttime or daytime attacks are more damaging would allow policymakers to better allocate resources when dealing with the treat of aerial attack.
+This study models the German bombing campaign of Britain during the Second World War by analyzing the data collected by the Home Security War Room from 1939 to 1945 that summarised air raid damage.
+I use a multiple linear regression model to estimate the impact of the time an attack occurred on the resulting casualties."
 format: pdf
@@
 
-I used the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+I used the statistical programming language R [@citeR] for data cleaning and analysis. The R packages used were
+
+.... Our data [@airraids].... Following @tellingstories, we consider...
 

### 2024-12-03 paper/paper.qmd
updated measurement

@@
 library(tidyverse)
-library(palmerpenguins)
 library(arrow)
@@
 
-I used the statistical programming language R [@citeR] for data cleaning and analysis. The R packages used were
+This study uses the [@airraids]
+
+I used the statistical programming language R [@citeR] for data cleaning and analysis.
+The R packages used were **`tidyverse`** for data manipulation and visualization [@tidyverse], **`here`** for file path management [@here], **`lubridate`** for date handling [@lubridate], **`arrow`** for data storage and processing [@arrow], **`testthat`** for unit testing [@testthat], **`readr`** for reading structured data [@readr], **`dplyr`** for data manipulation [@dplyr], **`stringr`** for string operations [@stringr], **`ggplot2`** for creating visualizations [@ggplot2], **`caret`** for modeling [@caret], **`kableExtra`** for creating tables [@kableExtra], and **`modelsummary`** for summarizing models and results [@modelsummary].
 
 .... Our data [@airraids].... Following @tellingstories, we consider...
+
+arrow /
+here / 
+readr /
+usethis
+dplyr / 
+testthat / 
+rstanarm
+ggplot2 /
+brms
+
+## Measurement
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The measurement process refers to the attempt to explain real world events through data, such as an individual act of warfare, as represented by the use of a self-propelled (missile) or gravity released explosive (bomb) to strike land or sea target, into numerical entries describing this event.
+Each entry in the dataset represents a bombing incident that took place in the UK during WW2.
 

### 2024-12-03 paper/paper.qmd
updated preamble added measurement, outcome

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Air raids on the UK: understanding the effect of time on casualties"
+subtitle: "How nighttime can influence the destructiveness of warfare"
 author: 
@@
 [https://github.com/dennisnetchitailo/STA304P3](https://github.com/dennisnetchitailo/STA304P3)."
-date: 2 December 2024
+date: 02 December 2024
 date-format: long
@@
 
-Overview text
-
-## Measurement
-
 The measurement process refers to the attempt to explain real world events through data, such as an individual act of warfare, as represented by the use of a self-propelled (missile) or gravity released explosive (bomb) to strike land or sea target, into numerical entries describing this event.
 Each entry in the dataset represents a bombing incident that took place in the UK during WW2.
+
+**Country**: The country where an incident occurred. This includes England, Scotland, Wales, Northern Ireland, and the Channel Islands. 
+**Location**: 
+**Civil Defense Region**: A region defined by the military for purposes of organization. The United Kingdom was divided into 13 regions. 
+**Coordinates: Longitude and Latitude**: The location of an incident as recorded by the mercator projection.  
+**Date: Start & End**: The date an incident started and ended at a specific location. 
+**Time** The time of day an incident occurred. Day is defined as being between 6AM and 6PM. Night is defined as being between 6PM and 6AM. 
+**Casualties: Killed, Injured, Total**: The number of people killed or injured due to an incident. Total casualties is the sum of the number killed and injured in the same incident.
+
+## Data Cleaning
+The two datasets that were used to make the single dataset that provided the data for the analysis were cleaned to ensure accuracy and consistency.
+Missing values were dealt with by removing them from the analysis.
+Variables were converted into logical data types.
+Unneeded columns were removed.
+The two cleaned datasets were unified into one.
 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+The outcome variable is **`total_casualties`**.
+This is the primary dependent variable that the model is designed to predict.
+It represents the number of people killed or injured by air raids.
+The model seeks to identify the variables affecting this number.
+
+@fig-expectancy visualizes the distribution of life expectancy from 2015 to 2020 as percentages, combining a histogram and a density plot.
+A large concentration of observations is observed range of 0% to 25%, reflecting a significant proportion of countries with lower life expectancy relative to others in the dataset.
+Another visible concentration is around 75% to 100%, indicating a group of countries with relatively high life expectancy.
+The blue density curve overlays the histogram, providing a smoothed representation of the distribution.
+Peaks in the curve correspond to high frequency within the histogram.
+The density plot emphasized the bimodal nature of the distribution with two distinct peaks: one around 12 as low life expectancy percentages; another peak around 87as high life expectancy percentage.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: fig-timesdistribution
+#| fig-cap: "Distribution of casualties in the UK caused by air raids throughout the war."
 #| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
+# Create the histogram plot
+ggplot(combined_data, aes(x = factor(start_month), y = total_casualties, fill = as.factor(start_year))) +
+  geom_bar(stat = "identity", position = "dodge") +
+  scale_x_discrete(name = "Month") +
+  scale_y_continuous(name = "Total Casualties") +
+  labs(title = "Total Casualties by Month (1939-1945)", fill = "Year") +
   theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 ```
-
-Talk more about it.
 
@@
 Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+
+The **predictor variables** are the factors that are assumed to influence the number of casualties.
+
+1. **Civil Defense Region ('civil_defense_region')**: is a factor because different military regions differed in their capabilities to prevent and to respond to air raids.
 

### 2024-12-03 paper/paper.qmd
Add large files via Git LFS

@@
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
 The outcome variable is **`total_casualties`**.
@@
 
-@fig-expectancy visualizes the distribution of life expectancy from 2015 to 2020 as percentages, combining a histogram and a density plot.
-A large concentration of observations is observed range of 0% to 25%, reflecting a significant proportion of countries with lower life expectancy relative to others in the dataset.
-Another visible concentration is around 75% to 100%, indicating a group of countries with relatively high life expectancy.
-The blue density curve overlays the histogram, providing a smoothed representation of the distribution.
-Peaks in the curve correspond to high frequency within the histogram.
-The density plot emphasized the bimodal nature of the distribution with two distinct peaks: one around 12 as low life expectancy percentages; another peak around 87as high life expectancy percentage.
+@fig-monthdistribution visualizes the distribution of casualties from 1939 to 1945 on a monthly basis in a histogram.
 
 ```{r}
-#| label: fig-timesdistribution
+#| label: fig-monthdistribution
 #| fig-cap: "Distribution of casualties in the UK caused by air raids throughout the war."
@@
 ```
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+A large number of casualties occurred in the last months of 1940 and the first months of 1941.
+This coincides with "The Blitz", the name given to the most intense period of air raids on the UK, following the defeat of France in June 1940.
+The number of casualties rises month to month following the defeat of the Allies in western Europe, as Germany could concentrate its airforce on bombing the United Kingdom.
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| label: fig-yeardistribution
+#| fig-cap: "Distribution of casualties in the UK caused by air raids throughout the war."
 #| echo: false
+#| eval: true
 #| warning: false
 #| message: false
+# Create the histogram plot
+ggplot(combined_data, aes(x = factor(start_year), y = total_casualties)) +
+  geom_bar(stat = "identity", fill = "steelblue") +
+  scale_x_discrete(name = "Year") +
+  scale_y_continuous(name = "Total Casualties") +
+  labs(title = "Total Casualties by Year (1939-1945)") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 0, hjust = 0.5))
+```
+@fig-yeardistribution shows that
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+summary_data
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
+```{r}
+#| label: fig-yeardistribution
+#| fig-cap: "Distribution of casualties in the UK caused by air raids throughout the war."
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+# Create the histogram plot
+# Create the dual-axis plot
+ggplot(combined_data, aes(x = factor(start_year))) +
+  geom_bar(aes(y = total_casualties), stat = "identity", fill = "steelblue", alpha = 0.7) +
+  geom_line(aes(y = total_incidents * (max(total_casualties) / max(total_incidents)), group = 1), 
+            color = "darkred", size = 1) +
+  geom_point(aes(y = total_incidents * (max(total_casualties) / max(total_incidents))), 
+             color = "darkred", size = 2) +
+  scale_y_continuous(
+    name = "Total Casualties",
+    sec.axis = sec_axis(~ . * (max(combined_data$total_incidents) / max(combined_data$total_casualties)), 
+                        name = "Number of Incidents")
+  ) +
+  labs(title = "Total Casualties and Number of Incidents by Year (1939-1945)", 
+       x = "Year") +
   theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+  theme(
+    axis.text.x = element_text(angle = 0, hjust = 0.5),
+    axis.title.y.right = element_text(color = "darkred"),
+    axis.text.y.right = element_text(color = "darkred")
+  )
 ```
@@
 
-1. **Civil Defense Region ('civil_defense_region')**: is a factor because different military regions differed in their capabilities to prevent and to respond to air raids.
+1.
+**Time ('time_binary')**.
+This variable represents whether an incident occurred during daytime, defined as between 6AM and 6PM, or nighttime, defined as between 6PM and 6AM.
+The assumption here is that the response time of medical, military, and police services differs at night, given more people at rest (implying worse response time), but also less traffic (implying faster response time).
+This is the key variable being assessed, given
+
+2.
+**Country ('country')** is a factor, because each country's cultural, social, and economic conditions can influence how much damage an air raid has by influencing factors such as the the abundance of underground shelter (such as the London Underground) or the quality of building construction (making a structure more or less susceptible to collapse).
+
+3.
+**Date ('start_year') and ('start_month')** is a factor because over time, technology and capabilities changed for both sides in the war.
+Equally, seasonal effects may have played a role.
+For instance, in wintertime, it is possible that there could have been cases of frostbite caused by the destruction of shelter and heating infrastructure.
+
+4.
+**Civil Defense Region ('civil_defense_region')**: is a factor because different military regions differed in their capabilities to prevent and to respond to air raids, as well as in intensity of targeting by the German airforce.
 
@@
 first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+  readRDS(file = here::here("STA304P3/models/refined_model.rds"))
 ```

### 2024-12-03 scripts/03-clean_data.R

@@ -148,6 +148,14 @@ cleaned_data <- combined_data %>%
                          ifelse(time == "Day", 0, NA))
   )
 
+# Remove rows with NA except for 'proportion_killed'
+cleaned_data <- combined_data %>%
+  filter(if_all(everything(), ~ !is.na(.) | cur_column() == "proportion_killed"))
+
+# Verify the result
+summary(cleaned_data)
+
+
 
 # View the cleaned dataset
 head(cleaned_data)

### 2024-12-03 scripts/05-exploratory_data_analysis.R

@@ -10,7 +10,7 @@
   # - The `rstanarm` package must be installed and loaded
   # - The `ggplot2` package must be installed and loaded
   # - The `brms` package must be installed and loaded
-
+  # - The `broom.mixed` package must be installed and loaded
 # Any other information needed? [...UPDATE THIS...]
 
 
@@ -20,7 +20,7 @@ library(rstanarm)
 library(dplyr)
 library(ggplot2)
 library(brms)
-
+library(broom.mixed)
 
 source_folder = "C:/Users/Dennis Netchitailo/Documents/STA304P3"
 
@@ -29,6 +29,15 @@ load_csv <- function(file_name, base_folder = source_folder) {
   data <- read.csv(file_path)
   return(data)
 }
+
+rtools_path <- "C:/RBuildTools/4.4"
+path_to_usr_bin <- file.path(rtools_path, "usr/bin")
+path_to_mingw_bin <- file.path(rtools_path, "mingw_64/bin")
+Sys.setenv(PATH = paste(Sys.getenv("PATH"), path_to_usr_bin, path_to_mingw_bin, sep = .Platform$path.sep))
+Sys.setenv(R_BUILD_TOOLS_PATH = "C:/RTools")
+Sys.which("make")
+
+
 
 #### Load Datasets ####
 
@@ -169,7 +178,9 @@ ggplot(combined_data, aes(x = as.factor(time_binary), y = total_casualties)) +
   labs(x = "Time (0 = Day, 1 = Night)", y = "Total Casualties") +
   theme_minimal()
 
+### Models ###
 
+options(mc.cores = parallel::detectCores())
 # Fit a Zero-Inflated Negative Binomial model
 zero_inflated_model <- brm(
   formula = total_casualties ~ time_binary + (1 | casualty_group),  # Add random effects if needed
@@ -204,18 +215,76 @@ ggplot(combined_data, aes(x = time_binary, y = total_casualties)) +
   theme_minimal()
 
 
+refined_model <- brm(
+  total_casualties ~ time_binary + severity + population_density + 
+    (1 | casualty_group) + (1 | civil_defense_region),
+  data = combined_data,
+  family = zero_inflated_negbinomial(),
+  prior = c(
+    prior(normal(0, 1), class = "b"),
+    prior(exponential(1), class = "sd")
+  ),
+  chains = 4,
+  cores = parallel::detectCores(),
+  iter = 2000
+)
+## New Model ___________________________________________________________________
 
[…]

### 2024-12-03 scripts/03-clean_data.R
cleaned up the file

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
+# Purpose: Cleans the raw bombings and casualties data recorded by the Ministry of Home Security
 # Author: Dennis Netchitailo
 # Date: November 29 2024
 # Contact: dennis.netchitailo@mail.utoronto.ca 
@@ -9,7 +9,7 @@
   # - The `tidyverse` package must be installed and loaded
   # - The `dplyr` package must be installed and loaded
   # - The `arrow` package must be installed and loaded
-# Any other information needed? [...UPDATE THIS...]
+# Any other information needed? No.
 
 #### Workspace setup ####
 library(tidyverse)
@@ -20,7 +20,7 @@ library(arrow)
 
 # Copy and insert the path to the folder
 #source_folder = "" #Insert path here
-source_folder = "C:/Users/Dennis Netchitailo/Documents/STA304P3"
+source_folder = "C:/Users/Dennis Netchitailo/Documents/air-raids"
 
 load_csv <- function(file_name, base_folder = source_folder) {
   file_path <- file.path(base_folder, file_name)
@@ -47,14 +47,6 @@ cleaned_casualties_data$proportion_killed <-
 # Handle cases where total_casualties is zero to avoid division by zero
 cleaned_casualties_data$proportion_killed[cleaned_casualties_data$total_casualties == 0] <- NA
 
-# Create dummy variable for Killed
-cleaned_casualties_data$has_killed <- ifelse(cleaned_casualties_data$killed > 0, 1, 0)
-
-# Create dummy variable for Injured
-cleaned_casualties_data$has_injured <- ifelse(cleaned_casualties_data$injured > 0, 1, 0)
-
-# Create dummy variable for Casualties
-cleaned_casualties_data$has_casualties <- ifelse(cleaned_casualties_data$total_casualties > 0, 1, 0)
 
 ### BOMBINGS Dataset ###
 
@@ -64,25 +56,17 @@ cleaned_bombings_data$bombing_id <- as.integer(cleaned_bombings_data$bombing_id)
 # Remove whitespace before/after 
 cleaned_bombings_data$location <- trimws(cleaned_bombings_data$location)
 
-# Duration of bombing in days 
-cleaned_bombings_data$duration_days <- 
-  as.Date(cleaned_bombings_data$end_date) - as.Date(cleaned_bombings_data$start_date) + 1
-
 # Year of occurrence
[…]

### 2024-12-03 scripts/04-test_analysis_data.R
removed unneeded tests added one test for new column lethality

@@ -56,64 +56,27 @@ test_that("country is of class character", {
   expect_type(combined_data$country, "character")
 })
 
-test_that("start_date is of class double", {
-  expect_type(combined_data$start_date, "double")
+
+test_that("year is of class double", {
+  expect_type(combined_data$year, "double")
 })
 
-test_that("end_date is of class Date", {
-  expect_s3_class(combined_data$end_date, "Date")
+test_that("year values are within the range 1939-1945, ignoring NAs", {
+  expect_true(all(combined_data$year >= 1939 & combined_data$year <= 1945, na.rm = TRUE))
 })
 
-test_that("time is of class character", {
-  expect_type(combined_data$time, "character")
+test_that("month is of class double", {
+  expect_type(combined_data$month, "double")
 })
 
-test_that("duration_days is of class double", {
-  expect_type(combined_data$duration_days, "double")
+test_that("month values are within the range 1-12, ignoring NAs", {
+  expect_true(all(combined_data$month >= 1 & combined_data$month <= 12, na.rm = TRUE))
 })
 
-test_that("start_year is of class double", {
-  expect_type(combined_data$start_year, "double")
-})
-
-test_that("start_year values are within the range 1939-1945, ignoring NAs", {
-  expect_true(all(combined_data$start_year >= 1939 & combined_data$start_year <= 1945, na.rm = TRUE))
-})
-
-test_that("start_month is of class double", {
-  expect_type(combined_data$start_month, "double")
-})
-
-test_that("start_month values are within the range 1-12, ignoring NAs", {
-  expect_true(all(combined_data$start_month >= 1 & combined_data$start_month <= 12, na.rm = TRUE))
-})
-
-test_that("time_unknown is of class double", {
-  expect_type(combined_data$time_unknown, "double")
-})
-
-#### Additional Tests ####
-
-# # Test proportion_killed is between 0 and 1
-# test_that("proportion_killed values are between 0 and 1, ignoring NAs", {
-#   expect_true(all(combined_data$proportion_killed >= 0 & combined_data$proportion_killed <= 1, na.rm = TRUE))
-# })
-
-test_that("proportion_killed values are between 0 and 1, ignoring NAs", {
-  # Ensure the proportion is within range for non-NA values
-  expect_true(all(combined_data$proportion_killed >= 0 & combined_data$proportion_killed <= 1, na.rm = TRUE))
-})
-
-
-# Test presence of boolean flags
-test_that("has_killed is of class double", {
[…]

### 2024-12-03 other/datasheet/datasheet_template.qmd
updated preamble added abstract

---
title: "Datasheet for 'air raids dataset'"
subtitle: ""
author: 
  - Dennis Netchitailo
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "The UK government required data about the effects of air raids in order to understand the impact on the population and the military. Germany's goal was to weaken the UK through the air enough to allow for an amphibious invasion. It was critical for the UK to understand the effectiveness of Germany's strategy of trying to degrade the UK's morale, economy, and war fighting capabilities."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to analyze the impact of air raids in the UK during WWII, focusing on casualties and incident patterns. It aimed to provide insights into the frequency and severity of incidents over time.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - TBD
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - TBD
4. *Any other comments?*
    - TBD

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- TBD
2. *How many instances are there in total (of each type, if appropriate)?*
	- TBD
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- TBD
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- TBD
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- TBD
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- TBD
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- TBD
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- TBD
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- TBD
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- TBD
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- TBD
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- TBD
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- TBD
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- TBD
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- TBD
16. *Any other comments?*
	- TBD

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- TBD
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- TBD
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- TBD
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- TBD
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- TBD
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- TBD
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- TBD
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- TBD
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- TBD
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
12. *Any other comments?*
	- TBD

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- TBD
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- TBD
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- TBD
4. *Any other comments?*
	- TBD

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- TBD
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- TBD
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- TBD
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- TBD

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- TBD
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- TBD
3. *When will the dataset be distributed?*
	- TBD
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- TBD
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- TBD
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- TBD
7. *Any other comments?*
	- TBD

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- TBD
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- TBD
3. *Is there an erratum? If so, please provide a link or other access point.*
	- TBD
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- TBD
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- TBD
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- TBD
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- TBD
8. *Any other comments?*
	- TBD

\newpage

# References

### 2024-12-03 scripts/03-clean_data.R
removed local path

@@ -20,7 +20,6 @@ library(arrow)
 
 # Copy and insert the path to the folder
 #source_folder = "" #Insert path here
-source_folder = "C:/Users/Dennis Netchitailo/Documents/air-raids"
 
 load_csv <- function(file_name, base_folder = source_folder) {
   file_path <- file.path(base_folder, file_name)

### 2024-12-03 scripts/05-exploratory_data_analysis.R
updated preamble removed local path

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
+# Purpose: Models the combined_data dataset 
 # Author: Dennis Netchitailo
 # Date: 30 November 2024
 # Contact: dennis.netchitailo@mail.utoronto.ca 
@@ -11,7 +11,7 @@
   # - The `ggplot2` package must be installed and loaded
   # - The `brms` package must be installed and loaded
   # - The `broom.mixed` package must be installed and loaded
-# Any other information needed? [...UPDATE THIS...]
+# Any other information needed? No.
 
 
 #### Workspace setup ####
@@ -22,7 +22,7 @@ library(ggplot2)
 library(brms)
 library(broom.mixed)
 
-source_folder = "C:/Users/Dennis Netchitailo/Documents/STA304P3"
+source_folder = "" # Insert local path here
 
 load_csv <- function(file_name, base_folder = source_folder) {
   file_path <- file.path(base_folder, file_name)

### 2024-12-04 other/datasheet/datasheet_template.qmd
updated datasheet

@@
 ---
-title: "Datasheet for 'air raids dataset'"
-subtitle: ""
+title: "Datasheet for the 'air raids dataset'"
+subtitle: "Supplementary document to give context to the data"
 author: 
   - Dennis Netchitailo
-thanks: "Code and data are available at: LINK."
-date: "`r Sys.time()`"
-date-format: "D MMMM YYYY"
+thanks: "Code and data are available at: https://github.com/dennisnetchitailo/air-raids."
+date: "03 December 2024"
+date-format: 
 abstract: "The UK government required data about the effects of air raids in order to understand the impact on the population and the military.
@@
 number-sections: true
-bibliography: references.bib
+bibliography: "C:/Users/Dennis Netchitailo/Documents/air-raids/paper/references.bib"
 ---
@@
 2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
-    - TBD
+    - The UK Ministry of Home Security as part of the Home Security Daily Intelligence Reports
 3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
-    - TBD
+    - The ministry was government funded. 
 4. *Any other comments?*
-    - TBD
+    - The dataset made up the vast amount of data collection that was undertaken by the UK government to aid the war effort.
 
@@
 Please provide a description.*
-	- TBD
+	- Each instance represents an air raid incident, including data on casualties, incidents, time, and location.
 2. *How many instances are there in total (of each type, if appropriate)?*
-	- TBD
+	- There are 32869 instances. 
 3.
@@
 If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
-	- TBD
+	- The dataset contains all possible instances.
+It is possible some instances were not recorded, but this would have had to have occurred in areas without human observation (uninhabited areas or the English channel), which would imply a lack of damage to people or infrastructure.
 4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
-	- TBD
+	- Processed features such as year, month, location, casualties, and lethality category.
 5. *Is there a label or target associated with each instance? If so, please provide a description.*
-	- TBD
+	- Yes, the lethality_category variable is used to classify incidents by severity.
 6.
@@
 This does not include intentionally removed information, but might include, for example, redacted text.*
-	- TBD
+	- Some data points such casualty numbers are unavailable due to incomplete records.
 7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
-	- TBD
+	- There is an explicit relationship between years, regions, and casualties.
 8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
-	- TBD
+	- There are no data splits used in this analysis. 
 9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
-	- TBD
+	- The errors would relate to incomplete historical records. It wasn't always possible to record a location or a casualty number due to complexity of determining how it is to be measured. 
 10.
@@
 Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
-	- TBD
+	- The dataset is self-contained, but links to the source historical archive for verification. 
 11.
@@
 If so, please provide a description.*
-	- TBD
+	- No, as the data describes a historical event in the aggregate. No personal details of any kind are in the data. 
 12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
@@
 If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
-	- TBD
+	- The dataset does not identify any sub-populations.
 14.
@@
 If so, please describe how.*
-	- TBD
+	- It is indirectly possible.
+With the coordinates, it is possible to identify current residents at places with incidents, but it would not be possible to identity the people at the incident.
+However, given another dataset that recorded the location from where a casualty was taken, it would be possible to link an individual to an incident.
 15.
@@
 If so, please provide a description.*
-	- TBD
+	- The sensitivity of the data is that a current location may be identified as having one been struck during an air raid.
+This could, for example, influence the sale of a building, if it became known that it had once been struck, due to a fear of its foundation being weakened.
 16. *Any other comments?*
-	- TBD
+	- The dataset analyzes contains high level data. Even the coordinate data, the most sensitive and identifying of all the data, doesn't cover all individual buildings that were struck.
 
@@
 If so, please describe how.*
-	- TBD
+	- The dataset was acquired from the datasheet made publicly available by Alex Cookson. 
+	- Likely from historical records, government archives, or academic sources. This would involve direct observation of targeted locations, recording 
 2.
@@
 How were these mechanisms or procedures validated?*
-	- TBD
+	- Manual digitization from historical documents.
 3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
-	- TBD
+	- The dataset is exhaustive. 
 4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
-	- TBD
+	- The data was collected by police, air raid wardens, and military personnel. This was part of their official duties. 
 5.
@@
 If not, please describe the timeframe in which the data associated with the instances was created.*
-	- TBD
+	- The data was collected from 1939 to 1945, which matches the timeframe of the data. 
 6.
@@
 If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- TBD
+	- No. 
 7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
-	- TBD
+	- Third-party government historical archives.
 8.
@@
 If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
-	- TBD
+	- Likely not. 
 9.
@@
 If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
-	- TBD
+	- Unlikely that consent was obtained. 
 10.
@@
 If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
-	- TBD
+	- See above. 
 11.
@@
 If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- TBD
+	- No. 
 12. *Any other comments?*
-	- TBD
+	-
 
@@
 If not, you may skip the remaining questions in this section.*
-	- TBD
+	- Data cleaning, creating derived variables like lethality_category, and handling missing values.
 2.
@@
 If so, please provide a link or other access point to the "raw" data.*
-	- TBD
+	- Yes, available at: https://github.com/tacookson/data/tree/master/britain-bombing-ww2/raw
 3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
-	- TBD
+	- Unknown what software was used. 
 4. *Any other comments?*
@@
 Is there anything a dataset consumer could do to mitigate these risks or harms?*
-	- TBD
+	- Be aware of the historical context of the dataset. Avoid misinterpretation. 
 5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
-	- TBD
+	- Avoid use in contexts that trivialize or sensationalize casualty data.
 6. *Any other comments?*
-	- TBD
+	- Be ethical and respectful in using this data for analysis or for interpetation.
 
@@
 If so, please provide a description.*
-	- TBD
+	- No.
 2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
-	- TBD
+	- The dataset is freely available on GitHub. 
 3. *When will the dataset be distributed?*
-	- TBD
+	- After review and validation.
 4.
@@
 If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
-	- TBD
+	- Open data license or public domain for historical data.
 5.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
-	- TBD
+	- No.
 6.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
-	- TBD
+	- No. 
 7. *Any other comments?*
-	- TBD
+	-
 
@@
 1. *Who will be supporting/hosting/maintaining the dataset?*
-	- TBD
+	- The author, Dennis Netchitailo.
 2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
-	- TBD
+	- By email: dennis.netchitailo@mail.utoronto.ca
 3. *Is there an erratum? If so, please provide a link or other access point.*
-	- TBD
+	- No.
 4.
@@
 If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
-	- TBD
+	- The dataset will only be updated to correct errors. 
 5.
@@
 If so, please describe these limits and explain how they will be enforced.*
-	- TBD
+	- Not applicable for historical data.
 6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
-	- TBD
+	- No.
 7.
@@
 If so, please provide a description.*
-	- TBD
+	- Contact by email.
 8. *Any other comments?*
-	- TBD
+	-
 

### 2024-12-04 paper/paper.qmd
updated paper

@@
 
+#| include: false
+setwd(here::here())  # Sets the working directory to the root of the project
+
+combined_data <- read.csv("data/02-analysis_data/combined_data.csv")
 ```
@@
 
-////Overview paragraph////
 On September 1st, 1939, the German armed forces crossed the Polish border. The United Kingdom declared war on Germany on September 3rd, 1939. 
@@
 Part of its responsibility was tracking data related to air raids.
-This data
+This data found more than 66 thousand people and 221 thousand people had been killed or injured by Germany's air campaign against the UK.
+Approximately 0.7% of the UK's 1939 population of 41.46 million would be killed or injured by air raids between 1939 to 1945.
+With 32 000 incidents, there were an average of 10 casualties per incident, including 2 killed and 8 injured.
+Yet, there is great variation.
+Some incidents resulted in hundreds of casualties.
+While many others resulted in zero casualties.
+Understanding what factors lead to such drastic differences, often ascribed to luck, is key.
 
-////Estimand paragraph////
 The estimand is the number of casualties sustained during an attack.
-I use the Home Security data with a multiple linear regression model to estimate the difference in casualties during nighttime air raids compared to daytime air raids
+I use the Home Security data with a multiple linear regression model to estimate the difference in casualties during nighttime air raids compared to daytime air raids.
 
-////Results paragraph////
+This paper's goal is to try to develop a model that predicts casualties based on the time of an air raid by controlling for multiple variables: country, military region, time, date, and the number of incidents occurring.
 
-////Why it matters paragraph////
+The analysis of this data can help countries facing war, especially as it relates to missile or airstrikes, better understand how to allocate resources to deal with casualties.
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The paper is structured as follows: @sec-data discusses the data and data cleaning process, followed by the analysis of key variables and relationships in @sec-model.
+@sec-results then presents the results from the model, followed by a discussion of the implications of these findings in @sec-discussion.
+@sec-discussion also concludes with suggestions for future research and practical applications of the model.
 
@@
 
-This study uses the [@airraids]
+This study uses the [@airraids] dataset to study the relationship between variables such as time to casualties.
 
 I used the statistical programming language R [@citeR] for data cleaning and analysis.
-The R packages used were **`tidyverse`** for data manipulation and visualization [@tidyverse], **`here`** for file path management [@here], **`lubridate`** for date handling [@lubridate], **`arrow`** for data storage and processing [@arrow], **`testthat`** for unit testing [@testthat], **`readr`** for reading structured data [@readr], **`dplyr`** for data manipulation [@dplyr], **`stringr`** for string operations [@stringr], **`ggplot2`** for creating visualizations [@ggplot2], **`caret`** for modeling [@caret], **`kableExtra`** for creating tables [@kableExtra], and **`modelsummary`** for summarizing models and results [@modelsummary].
-
-.... Our data [@airraids].... Following @tellingstories, we consider...
-
-arrow /
-here / 
-readr /
-usethis
-dplyr / 
-testthat / 
-rstanarm
-ggplot2 /
-brms
+The R packages used were **usethis** for file paths [@usethis], **`tidyverse`** for data manipulation and visualization [@tidyverse], **`here`** for file path management [@here], **`lubridate`** for date handling [@lubridate], **`arrow`** for data storage and processing [@arrow], **`testthat`** for unit testing [@testthat], **`readr`** for reading structured data [@readr], **`dplyr`** for data manipulation [@dplyr], **`stringr`** for string operations [@stringr], **`ggplot2`** for creating visualizations [@ggplot2], **`brms`** for modeling [@brms], and **`modelsummary`** for summarizing models and results [@modelsummary].
 
@@
 **Country**: The country where an incident occurred. This includes England, Scotland, Wales, Northern Ireland, and the Channel Islands. 
-**Location**: 
 **Civil Defense Region**: A region defined by the military for purposes of organization. The United Kingdom was divided into 13 regions. 
 **Coordinates: Longitude and Latitude**: The location of an incident as recorded by the mercator projection.  
-**Date: Start & End**: The date an incident started and ended at a specific location. 
+**Date: Year & Month**: The year & month an incident occurred at a specific location. 
 **Time** The time of day an incident occurred. Day is defined as being between 6AM and 6PM. Night is defined as being between 6PM and 6AM. 
 **Casualties: Killed, Injured, Total**: The number of people killed or injured due to an incident. Total casualties is the sum of the number killed and injured in the same incident.
+**Lethality**: the proportion of casualties that were killed.
 
@@
 # Create the histogram plot
-ggplot(combined_data, aes(x = factor(start_month), y = total_casualties, fill = as.factor(start_year))) +
+ggplot(combined_data, aes(x = factor(month), y = total_casualties, fill = as.factor(year))) +
   geom_bar(stat = "identity", position = "dodge") +
@@
 The number of casualties rises month to month following the defeat of the Allies in western Europe, as Germany could concentrate its airforce on bombing the United Kingdom.
+Based on this graph, a seasonal trend is not observed.
 
@@
 # Create the histogram plot
-ggplot(combined_data, aes(x = factor(start_year), y = total_casualties)) +
+ggplot(combined_data, aes(x = factor(year), y = total_casualties)) +
   geom_bar(stat = "identity", fill = "steelblue") +
@@
 ```
-@fig-yeardistribution shows that
-
-summary_data
+@fig-yeardistribution reflects the events going on in Europe at the time.
+In 1939, Germany occupied Poland.
+In 1940, it turned its attention to the west, and coinciding with the invasion of France, Germany's mass air raids against the UK began.
+This continued into 1941.
+In 1941, Germany invaded the Soviet Union, redirecting Germany's resources to the east.
+The renewed attention in 1944 coincides with the allied invasion of Normandy.
+Finally, in 1945, Germany surrended in May.
 
 ```{r}
-#| label: fig-yeardistribution
+#| label: fig-incidentyeardistribution
 #| fig-cap: "Distribution of casualties in the UK caused by air raids throughout the war."
@@
 #| message: false
-# Create the histogram plot
-# Create the dual-axis plot
-ggplot(combined_data, aes(x = factor(start_year))) +
-  geom_bar(aes(y = total_casualties), stat = "identity", fill = "steelblue", alpha = 0.7) +
-  geom_line(aes(y = total_incidents * (max(total_casualties) / max(total_incidents)), group = 1), 
-            color = "darkred", size = 1) +
-  geom_point(aes(y = total_incidents * (max(total_casualties) / max(total_incidents))), 
-             color = "darkred", size = 2) +
-  scale_y_continuous(
-    name = "Total Casualties",
-    sec.axis = sec_axis(~ . * (max(combined_data$total_incidents) / max(combined_data$total_casualties)), 
-                        name = "Number of Incidents")
-  ) +
-  labs(title = "Total Casualties and Number of Incidents by Year (1939-1945)", 
-       x = "Year") +
+# Create a bar plot showing incidents per year
+ggplot(combined_data, aes(x = factor(year))) +
+  geom_bar(aes(y = total_incidents), stat = "identity", fill = "skyblue", alpha = 0.7) +
+  labs(title = "Total Incidents per Year (1939-1945)", 
+       x = "Year", 
+       y = "Number of Incidents") +
   theme_minimal() +
-  theme(
-    axis.text.x = element_text(angle = 0, hjust = 0.5),
-    axis.title.y.right = element_text(color = "darkred"),
-    axis.text.y.right = element_text(color = "darkred")
-  )
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for clarity
+
 ```
-
-Talk way more about it.
+@fig-incidentyeardistribution reflects the events going on in Europe at the time.
+In 1939, Germany occupied Poland.
+In 1940, it turned its attention to the west, and coinciding with the invasion of France, Germany's mass air raids against the UK began.
+This continued into 1941.
+In 1941, Germany invaded the Soviet Union, redirecting Germany's resources to the east.
+The renewed attention in 1944 coincides with the allied invasion of Normandy.
+Finally, in 1945, Germany surrended in May.
+As @fig-yeardistribution showed, though there were more casualties in 1941 than in 1940, there were actually more incidents in 1940.
 
 ## Predictor variables
-
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
 
@@
 The assumption here is that the response time of medical, military, and police services differs at night, given more people at rest (implying worse response time), but also less traffic (implying faster response time).
-This is the key variable being assessed, given
+This is the key variable being assessed.
 
@@
 3.
-**Date ('start_year') and ('start_month')** is a factor because over time, technology and capabilities changed for both sides in the war.
+**Date ('year') and ('month')** is a factor because over time, technology and capabilities changed for both sides in the war.
 Equally, seasonal effects may have played a role.
@@
 
+5. **Casualties ('killed'), ('injured'), ('total_casualties'), and ('lethality_category')**: is a factor that assesses the number and types of casualties, as well as how deadly an air raid was.
+
 # Model
 
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The goal of this build a model using R [@citeR] to predict the difference in casualties caused in an incident based on several predictor variables.
 
@@
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+$$difference\_casualties_i​=\beta_0​+s(year_i​)+\beta_1​(month_i)+\beta_3​(time\_binary_i)+\beta_4​(lethality\_category_i)+\beta_5​(country_i)+\beta_6​(civil\_defense\_region_i)+\gamma_1​(1|casualty\_group_i)+\gamma_2​(1|civil\_defense\_region_i)+ ϵ_i​$$
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+I ran the model in R [@citeR] using the `brsm` package of @brms. I used the default priors from `brms`.
+
+Where:
+
+-   $difference\_casualties_i$ is the number of casualties for the i-th incident.
+
+-   $beta_0$ is the intercept term (baseline value when all predictors are zero).
+
+-   $year_i$ is the year the incident occurred in.
+
+-   $month_i$ is the month the incident occurred in.
+
+-   $time\_binary_i$  is whether the incident occurred during daytime or nighttime.
+
+-   $lethality\_category_i$​ is how deadly an air raid was.
+
+-   $country_i$​ is the country an incident occurred in.
+
+-   $civil\_defense\_region_i$​ is the military region an incident occurred in.
+
+-   $1|casualty\_group_i$​ is a random intercept term, accounting for variability caused by grouping.
+
+-   $1|civil\_defense\_region_i$​ is a random intercept term, accounting for variability caused by grouping.
+
+-   $\epsilon_i$ is the error term, assumed to be normally distributed with a mean of 0 and constant variance $ϵ_i​∼N(0,\sigma^2)$
 
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+The zero-inflated negative binomial distribution was chosen because of the nature of the data.
+First off, the response variable (total_casualties) is a count with non-negative integers.
+Secondly, due to the variance being larger than the mean, indicating overdispersion, the Poisson distribution would be innaprpriate, due to its assumption of the mean and variance being equal.
+The negative binomial distribution accounts for this overdispersion by introducing an additional parameter that models the variance separately from the mean, making it a more flexible choice for this data.
+Thirdly, relating to the large variance, but low mean, there are many zeros in the data.
+I.e.: many incidents had zero casualties.
+But a relatively small number of incidents have dozens, hundreds, or thousands of casualties.
+This is why the model was chosen to be a zero-inflated binomial distribution, to take into account the excess number of zeros.
+
+Other models such as a linear regression were considered, but deemed not to be suited to the characteristics of this dataset: data being a count, over-dispersion, excess zeros.
 
@@
 
-library(rstanarm)
+library(modelsummary)
+
+# Summarize the first model
+modelsummary(first_model, 
+             statistic = c("estimate", "std.error", "conf.low", "conf.high")
 
 first_model <-
-  readRDS(file = here::here("STA304P3/models/refined_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+  readRDS(file = here::here("air-raids/models/first_model.rds"))
 ```
@@
 
-Weaknesses and next steps should also be included.
-
 \newpage
@@
 # Appendix {-}
+
+# Idealized Survey
+Purpose: better understand the human and economic impact.
+
+Provide details regarding each (sub)question.
+
+Q1 - What was the location of air raid?
+
+(a) Country
+
+(b) Civil Defense region
+
+(c) Longitude and Latitude
+
+(d) Municipality
+
+Q2 - When did it occur?
+
+(a) What was the date of the first enemy strike on this location?
+
+(b) In what hour did this strike occur?
+
+Q3 - What casualties are associated with this strike?
+
+  (a) How many people were killed?
+
+  (b) If causes of death are known, list them, and the number associated with each cause.
+
+  (c) How many people were injured?
+
+  (d) If types of injury are known, list them, and the number associated with each cause.
+
+  (e) If known, what are the ages of the victims?
+
+  (f) If known, what is the sex of the victims?
+
+  (g) If known, how many of the casualties were civilians?
+
+  (h) If known, how many of the casualties were military personnel?
+
+Q4 - What was the economic impact?
+
+  (a) What objects was struck?
+
+  (b) If any, what damage was sustained by nearby objects?
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
 ## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
 

### 2024-12-04 scripts/06-model_data.R
updated file

@@ -1,33 +1,45 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
+# Purpose: Models the effect of time of an air raid on casualties.
 # Author: Dennis Netchitailo
 # Date: 30 November 2024
 # Contact: dennis.netchitailo@mail.utoronto.ca 
 # License: --
-# Pre-requisites: [...UPDATE THIS...]
+# Pre-requisites: --
   # - The `tidyverse` package must be installed and loaded
   # - The `rstanarm` package must be installed and loaded
-# Any other information needed? [...UPDATE THIS...]
+
+# Any other information needed? No.
 
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(dplyr)
+library(ggplot2)
+library(brms)
+library(broom.mixed)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+#model_data <- read_csv("data/analysis_data/combined_data")
+model_data <- read_csv("C:/Users/Dennis Netchitailo/Documents/air-raids/data/02-analysis_data/combined_data.csv")
+
+#"C:/Users/Dennis Netchitailo/Documents/air-raids/data/02-analysis_data/combined_data.csv"
 
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+first_model <- brm(
+  total_casualties ~ 
+    year + month + time_binary + time_binary + lethality_category + country + civil_defense_region + 
+    (1 | casualty_group) + (1 | civil_defense_region),
+  data = combined_data,
+  family = zero_inflated_negbinomial(),
+  prior = c(
+    prior(normal(0, 1), class = "b"),
+    prior(exponential(1), class = "sd")
+  ),
+  chains = 4,
+  cores = 14,  # Use 14 cores for faster computation
+  iter = 2000
+)
 
 
 #### Save model ####

### 2024-12-04 scripts/05-exploratory_data_analysis.R
updated file

@@ -205,37 +205,9 @@ pp_check(zero_inflated_model)
 # Generate predicted values
 combined_data$predicted <- predict(zero_inflated_model)[, "Estimate"]
 
-# Plot actual vs. predicted
-library(ggplot2)
-ggplot(combined_data, aes(x = time_binary, y = total_casualties)) +
-  geom_point(alpha = 0.4) +
-  geom_line(aes(y = predicted), color = "red", size = 1) +
-  labs(x = "Time (0 = Day, 1 = Night)", y = "Total Casualties",
-       title = "Zero-Inflated Model: Total Casualties vs. Time") +
-  theme_minimal()
 
+## Model ___________________________________________________________________
 
-refined_model <- brm(
-  total_casualties ~ time_binary + severity + population_density + 
-    (1 | casualty_group) + (1 | civil_defense_region),
-  data = combined_data,
-  family = zero_inflated_negbinomial(),
-  prior = c(
-    prior(normal(0, 1), class = "b"),
-    prior(exponential(1), class = "sd")
-  ),
-  chains = 4,
-  cores = parallel::detectCores(),
-  iter = 2000
-)
-## New Model ___________________________________________________________________
-
-## New Model ___________________________________________________________________
-## New Model ___________________________________________________________________
-
-## New Model ___________________________________________________________________
-
-## New Model ___________________________________________________________________
 library(brms)
 
 # Refined model using additional variables
@@ -259,10 +231,6 @@ saveRDS(
   file = "models/refined_model.rds"
 )
 
-## New Model ___________________________________________________________________
-## New Model ___________________________________________________________________
-## New Model ___________________________________________________________________
-## New Model ___________________________________________________________________
 ## New Model ___________________________________________________________________
 
 #Partial Effects

### 2024-12-04 paper/paper.qmd
updated

@@
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
 # Results
@@
 # Summarize the first model
+first_model <-
+  readRDS(file = here::here("models/first_model.rds"))
+
 modelsummary(first_model, 
-             statistic = c("estimate", "std.error", "conf.low", "conf.high")
+             statistic = c("estimate", "std.error", "conf.low", "conf.high"))
 
-first_model <-
-  readRDS(file = here::here("air-raids/models/first_model.rds"))
 ```

### 2024-12-04 scripts/06-model_data.R
updated preamble

@@ -7,6 +7,9 @@
 # Pre-requisites: --
   # - The `tidyverse` package must be installed and loaded
   # - The `rstanarm` package must be installed and loaded
+  # - The `ggplot2` package must be installed and loaded
+  # - The `brms` package must be installed and loaded
+  # - The `broom.mixed` package must be installed and loaded
 
 # Any other information needed? No.
 

### 2024-12-04 scripts/05-exploratory_data_analysis.R
cleaned

@@ -47,10 +47,6 @@ casualties_analysis_data <- load_csv("data/02-analysis_data/analysis_data_casual
 # Load bombings_data.csv
 bombings_analysis_data <- load_csv("data/02-analysis_data/analysis_data_bombings.csv")
 
-
-#### Read data ####
-#analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
-
 ### Model data ####
 
 ## CASUALTIES DATASET ##

### 2024-12-04 scripts/04-test_analysis_data.R
updated preamble

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
+# Purpose: Tests the variables to ensure the proper type and valid values
 # Author: Dennis Netchitailo
 # Date: 29 November 2024
 # Contact: dennis.netchitailo@mail.utoronto.ca 

### 2024-12-04 scripts/03-clean_data.R
cleaned

@@ -180,6 +180,3 @@ combined_data <- read_csv("data/02-analysis_data/combined_data.csv")
 write_parquet(cleaned_casualties_data, "data/02-analysis_data/analysis_data_casualties.parquet")
 write_parquet(cleaned_bombings_data, "data/02-analysis_data/analysis_data_bombings.parquet")
 write_parquet(combined_data, "data/02-analysis_data/combined_data.parquet")
-
-
-

### 2024-12-04 scripts/06-model_data.R
removed local path

@@ -23,10 +23,8 @@ library(brms)
 library(broom.mixed)
 
 #### Read data ####
-#model_data <- read_csv("data/analysis_data/combined_data")
-model_data <- read_csv("C:/Users/Dennis Netchitailo/Documents/air-raids/data/02-analysis_data/combined_data.csv")
+model_data <- read_csv("data/analysis_data/combined_data")
 
-#"C:/Users/Dennis Netchitailo/Documents/air-raids/data/02-analysis_data/combined_data.csv"
 
 ### Model data ####
 first_model <- brm(

### 2024-12-04 paper/paper.qmd
updated

@@
 
-This study uses the [@airraids] dataset to study the relationship between variables such as time to casualties.
+This study uses the [@britainbombing] dataset to study the relationship between variables such as time to casualties.
 
 I used the statistical programming language R [@citeR] for data cleaning and analysis.
-The R packages used were **usethis** for file paths [@usethis], **`tidyverse`** for data manipulation and visualization [@tidyverse], **`here`** for file path management [@here], **`lubridate`** for date handling [@lubridate], **`arrow`** for data storage and processing [@arrow], **`testthat`** for unit testing [@testthat], **`readr`** for reading structured data [@readr], **`dplyr`** for data manipulation [@dplyr], **`stringr`** for string operations [@stringr], **`ggplot2`** for creating visualizations [@ggplot2], **`brms`** for modeling [@brms], and **`modelsummary`** for summarizing models and results [@modelsummary].
+The R packages used were **usethis** for file paths [@usethis], **`tidyverse`** for data manipulation and visualization [@tidyverse], **`here`** for file path management [@here], , **`arrow`** for data storage and processing [@arrow], **`testthat`** for unit testing [@testthat], **`readr`** for reading structured data [@readr], **`dplyr`** for data manipulation [@dplyr], , **`ggplot2`** for creating visualizations [@ggplot2], **`brms`** for modeling [@brms], and **`modelsummary`** for summarizing models and results [@modelsummary].
 
@@
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(modelsummary)
-
-# Summarize the first model
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-
-modelsummary(first_model, 
-             statistic = c("estimate", "std.error", "conf.low", "conf.high"))
-
-```
-
 # Discussion
@@
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
 ## Second discussion point
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
 
